---
title: "Drift Update"
author: "Lisa Vance"
date: "2026-05-04"
output: html_document
---

Update to YBFMP Drift Dataset encompassing 1998-2024

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 10000)
```

## Setup 

1.Setup and Download Libraries

```{r setup, results=FALSE, warning = FALSE, message = FALSE}
rm(list=ls(all=TRUE))

library(tidyverse)
library(gridExtra) # combine plots
library(stringr)
library(readr)
library(lubridate)
library(plotly)
library(viridis)
library(kableExtra)
library(tidylog)

```

2.Load Data, Check Variable Types, make initial changes

```{r load, results = FALSE, warning = FALSE, message = FALSE}
# Load data - several tables
phys <- read_csv("drift data/LT_phys_qc_20240118.csv")
phys2 <- read_csv("R_write/LT_phys_qc_20260512.csv")
catch <- read_csv("drift data/2019 and prior/Drift_Catch_20200918.csv")
catch2 <- read_csv("drift data/2019 and prior/Drift_Catch_Contractor_20210330.csv")
###########
catch3 <- read_csv("drift data/Drift_Catch_20260429.csv", skip=1)
catch4<- read_csv("drift data/Drift_20231030-20241231.csv")
##########
samp <- read_csv("drift data/2019 and prior/Drift_Samplings_20200918.csv")
samp2 <- read_csv("drift data/2019 and prior/Drift_Sampling_Access_20210330.csv")
#########
samp3 <- read_csv("drift data/DriftSampExcelData_20260429.csv", skip=1)
#########
tax <- read_csv("drift data/DriftTaxonomy.csv")
wy <- read_csv("WaterYearType_CDEC.csv") 
inundation <- read_csv("Yolo_Bypass_Inundation_1998-2024.csv")
#inv_tax_2 <- read_csv("Data/TblInvertsLookUpV2.csv")

# Add and change date formats
str(phys)
str(phys2)
# phys files seem to be in correct format post qaqc

str(catch)
str(catch2)
str(catch3)

catch$Time <- as.POSIXct(catch$Time,
                         format = "%m/%d/%Y %H:%M")

catch2$Date <- as.Date(catch2$Date, "%m/%d/%Y")
catch2$Time <- hms::as_hms(catch2$Time)
catch2$Datetime = paste(catch2$Date, catch2$Time)
catch2$Datetime <- as.POSIXct(catch2$Datetime, 
                           format = "%Y-%m-%d %H:%M:%S")

catch3 <- catch3 %>%
  filter(!(is.na(`Measuring program short name`))) %>%
  rename(Count = `Value...11`,
         LifeStage = `Value...12`,
         TaxonName = Observable,
         Date = `Sampling Event Date`,
         Time = `Sampling Event Time`,
         Station = `Sampling Area Number`,
         SAMCode = `Sampling Event Number`,
         SampleID = `Sample ID`,
         LabComments = lab_comments) %>%
  select(-c(`Measuring program short name`, `Observation Type Short Name`,
            SAMCode))

catch3$Date <- as.Date(catch3$Date, "%m/%d/%Y")
catch3$Time <- hms::as_hms(catch3$Time)
catch3$Datetime = paste(catch3$Date, catch3$Time)
catch3$Datetime <- as.POSIXct(catch3$Datetime, 
                           format = "%Y-%m-%d %H:%M:%S")

catch4 <- catch4 %>% 
  select(-c(measuring_program, sam_code, short_name)) %>% 
  rename(Date = date_col,
         Time = time,
         Station = station,
         SampleID = sample_id,
         Attribute = attribute,
         TaxonName = observable,
         Category = category,
         Count = count,
         LifeStage = life_stage,
         LabComments = lab_comments)

catch4$Date <- as.Date(catch4$Date, "%m/%d/%Y")
catch4$Time <- hms::as_hms(catch4$Time)
catch4$Datetime = paste(catch4$Date, catch4$Time)
catch4$Datetime <- as.POSIXct(catch4$Datetime, 
                           format = "%Y-%m-%d %H:%M:%S")
str(catch3)
str(catch4)

catch3ed <- bind_rows(catch3, catch4)

str(samp)
str(samp2)
str(samp3)

samp$Date<-as.Date(samp$Date,"%m/%d/%Y")
samp$`Date/Time` = as.POSIXct(samp$`Date/Time`, 
                           format = "%m/%d/%Y %H:%M:%S")

samp3 <- samp3 %>%
  filter(!(is.na(`Measuring program short name`))) %>%
  rename(FlowMeterStart = `Flow Meter Start`,
         FlowMeterEnd = `Flow Meter End`,
         Date = `Sampling Event Date`,
         Time = `Sampling Event Time`,
         Station = `Sampling Area Number`,
         SAMCode = `Sampling Event Number`,
         MeshSize = `Observation Area Name`,
         SampleAltered = `Sampling Altered`,
         ConditionCode = `Condition Code`,
         SetTime = `Set Time`,
         FlowMeterSpeed = `Flow Meter Speed`, 
         FieldComments = `Field Comments`,
         SlideCount = `Slide Count`,
         SampleVolume = `Sample Volume`,
         SubsampleNumber = `Subsample Number`,
         DilutionVolume = `Dilution Volume`,
         PhysicalDataID = `Physical Data ID`) %>%
  select(-c(`Observation Area Number`, `Spot Code (original/duplicate)`, `...27`, `...28`,
            `...29`, `...30`, `...31`, `...32`,
            `Spot Number`, `Flow Meter Start (50)`, `Flow Meter End (50)`,
            `Entered by`, `QAQC'd by`, `...6`, `Measuring program short name`,
            SAMCode))

samp3$Date <- as.Date(samp3$Date,"%m/%d/%Y")
samp3$Time <- hms::as_hms(samp3$Time)
samp3$Datetime = paste(samp3$Date, samp3$Time)
samp3$Datetime <- as.POSIXct(samp3$Datetime, 
                           format = "%Y-%m-%d %H:%M:%S")
inundation <- inundation %>%
  rename(Date = Dates)
inundation$Date<-ymd(inundation$Date)

str(phys)
str(catch)
```

3.Rename variables
4. Make changes already identified

```{r filter, results = FALSE, message = FALSE, warning = FALSE}


catch <- catch %>%
  mutate(SamplingID = str_extract(`Sampling number`, "[^/]+")) %>%
  rename(Datetime = Time,
         TaxonName = Observable,
         LifeStage = `Life Stage`) %>%
  select(-c(Value, `Sampling number`))
catch$SamplingID = as.numeric(catch$SamplingID)

samp <- samp %>% 
  rename(SamplingID = `Sampling number`,
         Datetime = `Date/Time`,
         InvertDataID = `Invertebrate Data ID`,
         ConditionCode = `Condition Code`,
         SetTime = `Set Time`,
         FlowMeterSpeed = `Flow Meter Speed`,
         FlowMeterStart = `Flow Meter Start`,
         FlowMeterEnd = `Flow Meter End`,
         LabComments = `Lab Comments`,
         FieldComments = `Field Comments`) %>%
  select(-`Invertebrate Catch ID`) %>% 
  arrange(Datetime)
samp$SamplingID <- as.numeric(samp$SamplingID)
summary(samp)

catch2 <- catch2 %>%
  rename(Count = Individuals,
         Category = `Insect/Non-Insect`,
         LifeStage = Stage,
         TaxonName = Name) %>%
  select(-c("Number", "Time"))

samp2 <- samp2 %>%
  rename(FlowMeterStart = DriftStartMeter,
         FlowMeterEnd = DriftEndMeter) %>%
  select(-c("StartTime", "StopTime"))




#check data types again
str(phys)
str(catch)
str(samp)
str(samp2)
str(catch2)
str(inundation)

```

5. Merge data tables, calculate Flowmeter difference, rearrange variables

```{r merge, message = FALSE, warning = FALSE}

samp_catch <- left_join(samp, catch, by = c("SamplingID", "Station")) %>%
  rename(Datetime = Datetime.x) %>%
  select(-c(Datetime.y))

# Merge physical data
samp_catch_phys0 <- left_join(phys, samp_catch, by = "PhysicalDataID") %>%
  rename(Station = Station.x,
         Date = Date.x,
         Time = Time.x,
         Datetime = Datetime.x) %>%
  select(-c(Station.y, Date.y, Time.y, Datetime.y)) %>%
 filter(!is.na(Station)) %>%
  select(-c("ConditionCode.x", "MeterSetTime", "FlowMeterStart.x", "FlowMeterEnd.x",
            "FlowMeterSpeed.x", "FieldComments.x")) %>%
  rename(FlowMeterStart = "FlowMeterStart.y",
         FlowMeterEnd = "FlowMeterEnd.y",
         FlowMeterSpeed = "FlowMeterSpeed.y",
         ConditionCode = "ConditionCode.y",
         FieldComments = "FieldComments.y") %>% 
  filter(Date < "2019-04-22") 
notjoinedPhysDataID <- anti_join(phys, samp_catch, by = "PhysicalDataID")

# For second part 2019, merge phys-samp, then add catch.
# For the additional data
phys_samp <- left_join(phys, samp2, by = c("PhysicalDataID")) %>%
  filter(Date > "2019-04-16" & Date < "2020-01-01") %>%
  mutate(SamplingID = "N/A") %>% 
    select(-c("ConditionCode.x", "MeterSetTime", "FlowMeterStart.x", "FlowMeterEnd.x",
            "FlowMeterSpeed.x", "FieldComments.x", "EnteredBy", "QA/QC'dBy")) %>%
  rename(FlowMeterStart = "FlowMeterStart.y",
         FlowMeterEnd = "FlowMeterEnd.y",
         FlowMeterSpeed = "FlowMeterSpeed.y",
         ConditionCode = "ConditionCode.y",
         FieldComments = "FieldComments.y") %>% 
  arrange(Datetime)

str(phys_samp)
str(catch2)


phys_samp_catch0 <- left_join(phys_samp, catch2, by = c("Date", "Station")) %>%
    select(c(event_id:Comment_PQC, SamplingID, InvertDataID:SetTime, FlowMeterSpeed, FlowMeterStart, FlowMeterEnd, LabComments, FieldComments, TaxonName, Count, Category, LifeStage)) %>% 
  rename(Datetime = Datetime.x)

str(phys_samp_catch0)

########group of 2020 on
phys3 <- bind_rows(phys, phys2) %>% 
  filter(year(Date) > 2019) %>% 
  arrange(Datetime)
#check january 2020 - in both access and excel
samp2ed <- samp2 %>% 
  filter(PhysicalDataID == 2010 | PhysicalDataID == 2005 |
           PhysicalDataID == 2009| PhysicalDataID == 2007)

phys3ed <- phys3 %>% 
  filter(MonthAbb == "Jan" & Year == 2020)

overlap <- left_join(phys3ed, samp2ed, by = "PhysicalDataID") %>% 
  select(-c(ConditionCode.x, MeterSetTime, FlowMeterStart.x, FlowMeterEnd.x, FlowMeterSpeed.x, FieldComments.x,
           EnteredBy, "QA/QC'dBy")) %>% 
  rename(FlowMeterStart = FlowMeterStart.y,
         FlowMeterEnd = FlowMeterEnd.y,
         FieldComments = FieldComments.y,
         ConditionCode = ConditionCode.y,
         FlowMeterSpeed = FlowMeterSpeed.y)

overlapcatch <- left_join(overlap, catch3ed, by = c("Date", "Time", "Station")) %>% 
  select(-(Datetime.x)) %>% 
  rename(Datetime = Datetime.y)

##########
phys3ed2 <- phys3 %>% 
  filter(!(MonthAbb == "Jan" & Year == 2020)) %>% 
  unique()

phys_samp3 <- left_join(phys3ed2, samp3, by = c("Date", "Time", "Station", 
                                                "ConditionCode", "MeshSize", "Datetime", "PhysicalDataID")) %>% 
  select(-c(SetTime, SlideCount, FlowMeterStart.y, FlowMeterEnd.y, FlowMeterSpeed.y, FieldComments.y)) %>% 
  rename(SetTime = MeterSetTime,
         FlowMeterStart = FlowMeterStart.x,
         FlowMeterEnd = FlowMeterEnd.x,
         FlowMeterSpeed = FlowMeterSpeed.x,
         FieldComments = FieldComments.x)

phys_samp_catch3 <- left_join(phys_samp3, catch3ed, by = c("Date", "Time", "Station")) %>% 
  select(-(Datetime.y)) %>% 
  rename(Datetime = Datetime.x)

sampcatchphysMerge <- rbind(samp_catch_phys0, phys_samp_catch0) %>% 
  arrange(Datetime)

sampcatchphysMerge2 <- bind_rows(overlapcatch, phys_samp_catch3)

sampcatchphysMergefinal <- bind_rows(sampcatchphysMerge, sampcatchphysMerge2) %>% 
  arrange(Datetime)

check <- sampcatchphysMergefinal %>% 
  filter(is.na(FlowMeterStart))

#write to csv total merge file
write_csv(sampcatchphysMergefinal, paste("R_write/Sampling_Drift_1998-2024.csv"))

# All samplings - remove catch info and find unique entries
sampUnique <- sampcatchphysMergefinal %>%
  select(-c(WeatherCode:Turbidity, Year:Comment_PQC,  LabComments, TaxonName:LifeStage)) %>%
  unique() %>%
  arrange(Datetime)
```


```{r}

# In the future, figure out what to do with lab comments, which for drift are at the observable level. If they never include lab comments relative to sample quality, leave these out. If they are included, need to add !is.na(LabComments to the SamplingQAQC command.)
sampA <- sampcatchphysMergefinal %>%
  select(-c(WeatherCode:Turbidity, Year:Comment_PQC,  TaxonName:LifeStage)) %>%
  unique() %>%
  arrange(Datetime)


```

Add Water Year and Water Year "Class" (W = W, D/C = D, AN/BN = A), calculate Flowdiff
```{r}
samp_catch_physMergefinal <- sampcatchphysMergefinal %>%
  mutate(WY = ifelse(Month >9, Year + 1, Year)) %>%
  left_join(wy, by = "WY") %>%
  select(-c(Index, WYType)) %>%
  mutate(Flowdiff = abs(FlowMeterEnd-FlowMeterStart))

```

## QAQC 

* QC Flags: 0 or Blank = no issues/1 = Comment but deemed Acceptable/ 2 = Suspect/ 3 = Highly Suspect
* Flag_SAMP: Sampling issues
* Flag_LAB: Lab issues
* Flag_FM: Flowmeter
* Flag_CPUE: CPUE

### QAQC Sampling and Lab issues

6. Export sampling table
7. Look at field sampling comments
8. Flag comments reflecting poor data quality (Flag_SAMP)
9. Add comment about what was flagged
    * FMMISSING if lacking flowmeter values
    * SAMP for other sampling issues 
    * Leave blank if no issues
10. Look at lab comments and flag if poor data quality
11. Add Comment_LAB
    * LAB if lab issues (e.g. unable to meet meso or microtally)
    * Leave blank if no issues

* This part is manual. Think about how to make this codeable, change our condition code flags or add a flag as part of data entry? 

```{r}
SamplingQAQC <- filter(sampUnique, !is.na(FieldComments) | ConditionCode>1)
SamplingQAQC$Flag_SAMP <-  ""
SamplingQAQC$Comment_SAMP <-""
SamplingQAQC$Flag_LAB <- ""
SamplingQAQC$Comment_LAB <- ""
today <- today()

SamplingQAQC1 <- SamplingQAQC %>% 
  filter(year(Date)>2022)
write.csv(SamplingQAQC1, paste("R_write/Drift/SamplingQAQC_", today, ".csv"))
SamplingQAQC2 <- SamplingQAQC %>% 
  filter(year(Date)<2023)
write.csv(SamplingQAQC2, paste("R_write/Drift/SamplingQAQC2_", today, ".csv"))
```

12. Read edited file in

```{r}
#read in QAQC'd data and bind previously done QAQC with new QAQC

SamplingQAQC_fill <- read.csv("R_write/Drift/SamplingQAQC2_Notated_20260512.csv")
SamplingQAQC_fill2 <- read.csv("R_write/Drift/SamplingQAQC_Notated_20260512.csv")
SamplingQAQC_fill_s <- SamplingQAQC_fill %>%
  select(c(event_id, PhysicalDataID, Flag_SAMP, Comment_SAMP, Flag_LAB, Comment_LAB))
SamplingQAQC_fill_s2 <- SamplingQAQC_fill2 %>%
  select(c(event_id, PhysicalDataID, Flag_SAMP, Comment_SAMP, Flag_LAB, Comment_LAB))

SamplingQAQC_fill_sfinal <- bind_rows(SamplingQAQC_fill_s, SamplingQAQC_fill_s2)
```
#in the future, utilize the QAQC file moving forward and bind to only the most recent updating data - reduce redundancy of work being done - code will need to be re-written to accommodate this in the future

### QAQC Flowmeters

13. Inundation data

    This data comes from Water Data Library, Stations at Fremont Weir (A02170) and I Street Bridge . If Stage Height > 32ft at Fremont Weir, Inundation = TRUE. If there is no data at Fremont Weir, and Stage Height > 17 ft at I Street Bridge, Inundation = TRUE. There was also 2 days of lag time added in for inundation (variable Inundation2). 
    
* Inundation used to QA/QC STTD data. 

```{r}
inundation <- inundation %>%
  mutate(Month = month(Date),
         Year = year(Date),
         WY = ifelse(Month > 9, Year + 1, Year))

# Modify sample table to include Flowdiff
samp3 <-sampUnique %>%
  mutate(Flowdiff = abs(FlowMeterEnd-FlowMeterStart),
         Flowdiff_s = Flowdiff/SetTime) 

# merged inundation and sampling
inundation_flow <- left_join(samp3, inundation, by = "Date") %>%
    filter(Flowdiff<100000) %>%
  filter(Station == "STTD") %>%
  mutate(Flowdiff_s = Flowdiff/SetTime,
         Inundation_n = ifelse(Inundation == "TRUE", 40000, 0))

# all inundation
inundation2 <- inundation %>%
  mutate(Inundation_n = ifelse(Inundation == "TRUE", 40000, 0),
         Inundation_n_low = ifelse(Inundation == "TRUE", 5000,0))

# Plot - only inundation events that correspond with sampling date
inplot1 <- ggplot(inundation_flow) + 
  geom_point(aes(x = Date, y = Flowdiff)) + 
  geom_col(aes(x = Date, y = Inundation_n), fill = "blue", linewidth = 2) + 
  labs(title = "Flowdiff with Inundation events corresponding with sample dates") +
  facet_grid(Station~.) +
  theme_bw()
ggplotly(inplot1)

# Plot - All inundation events
inplot2 <- ggplot() +
  geom_point(data = inundation_flow, aes(x = Date, y = Flowdiff)) +
  labs(title = "Flowdiff with all Inundation") +
  geom_col(data = inundation2, aes(x = Date, y = Inundation_n), fill = 
                 "blue", alpha = 0.6) + theme_bw()

inplot3 <- ggplot() +
  geom_point(data = inundation_flow, aes(x = Date, y = Flowdiff_s)) +
  labs(title = "Standardized Flowdiff with all Inundation") +
  geom_col(data = inundation2, aes(x = Date, y = Inundation_n_low), fill = 
                 "blue", alpha = 0.6) + theme_bw()

grid.arrange(inplot1, inplot2, inplot3, nrow = 3)


# Allow 10 days after last "inundation=TRUE" to also count as inundated. Merge with sample data. 
inundation4 <- inundation %>%
  mutate(Inundation2 = ifelse(lead(Inundation, 10) == "TRUE", "TRUE", Inundation)) %>%
  select(c(Date, Month:Inundation2))

samp_catch_phys <- left_join(samp_catch_physMergefinal, inundation4)%>%
  select(-SampleID)


# checkflow <- samp_catch_phys %>% 
#   filter(is.na(Flowdiff))
```

14. Look at distribution of flowmeter values

```{r}
# Histogram of values
samp_catch_phys$Month <- ordered(samp_catch_phys$Month)
FlowHist <- ggplot(samp_catch_phys, aes(Flowdiff)) + geom_histogram() +
  facet_wrap(~Station) + theme_bw()

# SHR vs STTD boxplot
FlowBox <- ggplot(samp_catch_phys) + geom_boxplot(aes(x = Station, y = Flowdiff)) + theme_bw()
ggplotly(FlowBox)


# SHR vs STTD boxplot by month
FlowBoxMonth <- ggplot(samp_catch_phys) + geom_boxplot(aes(x = Month, y = Flowdiff, fill = Month)) + facet_wrap(~Station) + scale_fill_viridis(discrete = TRUE) + theme_bw() 

# SHR vs STTD boxplot by year
FlowBoxYear <- ggplot(samp_catch_phys) + geom_boxplot(aes(x = ordered(WY), y = Flowdiff, fill = ordered(WY))) + facet_wrap(~Station) + theme_bw() 

# SetTime vs Flowdiff by station and month
FlowPoint <- ggplot(samp_catch_phys, aes(x = SetTime, y = Flowdiff, color = Month)) + geom_point(size = 3) + facet_wrap(~Station) + theme_bw()

# SetTime vs Flowdiff by station and flowmeter type
FlowPoint2 <- ggplot(samp_catch_phys, aes(x = SetTime, y = Flowdiff, color = FlowMeterSpeed)) + geom_point(size = 3) + facet_wrap(~Station) + theme_bw() + scale_color_viridis(discrete = TRUE) 


grid.arrange(FlowBox, FlowHist)
grid.arrange(FlowBoxMonth, FlowBoxYear)
grid.arrange(FlowPoint, FlowPoint2)

```

15. Change obvious flowmeter values
```{r QAQC-Flowmeters}
# Change known incorrect values
# Change FlowMeterStart and FlowMeterEnd, Recalculate Flowdiff
# SamplingID = 47736 Switched FlowMeterStart and FlowMeterEnd 
check <- samp_catch_phys %>% filter(Flowdiff < 0)
checklarge <- samp_catch_phys %>% filter(Flowdiff > 90000 & SetTime < 30)

samp_catch_phys$FlowMeterStart[samp_catch_phys$SamplingID==47736] <- 573926
samp_catch_phys$FlowMeterEnd[samp_catch_phys$SamplingID==47736] <- 581200
samp_catch_phys$Flowdiff[samp_catch_phys$SamplingID==47736] <- 581200-573926

samp_catch_phys$FlowMeterStart[samp_catch_phys$InvertDataID==978] <- 788767
samp_catch_phys$FlowMeterEnd[samp_catch_phys$InvertDataID==978] <- 	795100
samp_catch_phys$Flowdiff[samp_catch_phys$InvertDataID==978] <- 795100-788767

samp_catch_phys$Flowdiff[samp_catch_phys$event_id== "SHR_2022-09-06 07:29:00"] <- abs((999999-samp_catch_phys$FlowMeterStart[samp_catch_phys$event_id== "SHR_2022-09-06 07:29:00"]) + samp_catch_phys$FlowMeterEnd[samp_catch_phys$event_id== "SHR_2022-09-06 07:29:00"])

samp_catch_phys$Flowdiff[samp_catch_phys$event_id== "SHR_2023-05-02 09:37:00"] <- abs((999999-samp_catch_phys$FlowMeterStart[samp_catch_phys$event_id== "SHR_2023-05-02 09:37:00"]) - samp_catch_phys$FlowMeterEnd[samp_catch_phys$event_id== "SHR_2023-05-02 09:37:00"])

samp_catch_phys$Flowdiff[samp_catch_phys$event_id== "STTD_2023-06-26 12:24:00"] <- abs((999999-samp_catch_phys$FlowMeterStart[samp_catch_phys$event_id== "STTD_2023-06-26 12:24:00"]) - samp_catch_phys$FlowMeterEnd[samp_catch_phys$event_id== "STTD_2023-06-26 12:24:00"])

samp3$FlowMeterStart[samp3$SamplingID==47736] <- 573926
samp3$FlowMeterEnd[samp3$SamplingID==47736] <- 581200
samp3$Flowdiff[samp3$SamplingID==47736] <- 581200-573926

samp3$FlowMeterStart[samp3$InvertDataID==978] <- 788767
samp3$FlowMeterEnd[samp3$InvertDataID==978] <- 	795100
samp3$Flowdiff[samp3$InvertDataID==978] <- 795100-788767

samp3$Flowdiff[samp3$event_id== "SHR_2022-09-06 07:29:00"] <- abs((999999-samp3$FlowMeterStart[samp3$event_id== "SHR_2022-09-06 07:29:00"]) + samp3$FlowMeterEnd[samp3$event_id== "SHR_2022-09-06 07:29:00"])

samp3$Flowdiff[samp3$event_id== "SHR_2023-05-02 09:37:00"] <- abs((999999-samp3$FlowMeterEnd[samp3$event_id== "SHR_2023-05-02 09:37:00"]) - samp3$FlowMeterStart[samp3$event_id== "SHR_2023-05-02 09:37:00"])

samp3$Flowdiff[samp3$event_id== "STTD_2023-06-26 12:24:00"] <- abs((999999-samp3$FlowMeterStart[samp3$event_id== "STTD_2023-06-26 12:24:00"]) - samp3$FlowMeterEnd[samp3$event_id== "STTD_2023-06-26 12:24:00"])

```

16. Merge data with lab/sampling QAQC 
    * Replace NA with 0
```{r}
FM_Samp <- left_join(samp_catch_phys, SamplingQAQC_fill_sfinal, by = "event_id") %>%
  mutate(Flag_SAMP = replace(Flag_SAMP, is.na(Flag_SAMP), 0 ),
         Comment_SAMP = replace(Comment_SAMP, is.na(Comment_SAMP), 0),
         Flag_LAB = replace(Flag_LAB, is.na(Flag_LAB), 0),
         Comment_LAB = replace(Comment_LAB, is.na(Comment_LAB), 0)) %>%
  select(-c(PhysicalDataID.y))%>%
  rename(PhysicalDataID = "PhysicalDataID.x") %>%
  mutate(FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Regular", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Reg", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "REG", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Low", "LOW")) %>% 
  unique()
```


#these were notated before and need to be added in to the notated file - this is a temporary portion that can be removed for next publishing
```{r}
FM_Samp$Flag_SAMP[FM_Samp$event_id == "SHR_2001-04-04 14:20:00"] <- 3
FM_Samp$Comment_SAMP[FM_Samp$event_id == "SHR_2001-04-04 14:20:00"] <- "FM"

FM_Samp$Flag_SAMP[FM_Samp$event_id == "SHR_2013-03-20 08:56:00"] <- 3
FM_Samp$Comment_SAMP[FM_Samp$event_id == "SHR_2013-03-20 08:56:00"] <- "SAMP"

FM_Samp$Flag_LAB[FM_Samp$event_id == "SHR_2013-06-12 09:30:00"] <- 3
FM_Samp$Comment_LAB[FM_Samp$event_id == "SHR_2013-06-12 09:30:00"] <- "LAB"

FM_Samp$Flag_LAB[FM_Samp$event_id == "STTD_2013-07-11 10:10:00"] <- 3
FM_Samp$Comment_LAB[FM_Samp$event_id == "STTD_2013-07-11 10:10:00"] <- "LAB"

FM_Samp$Flag_LAB[FM_Samp$event_id == "SHR_2013-08-23 11:55:00"] <- 3
FM_Samp$Comment_LAB[FM_Samp$event_id == "SHR_2013-08-23 11:55:00"] <- "LAB"

FM_Samp$Flag_SAMP[FM_Samp$event_id == "SHR_2013-10-02 09:20:00"] <- 3
FM_Samp$Comment_SAMP[FM_Samp$event_id == "SHR_2013-10-02 09:20:00"] <- "SAMP"

FM_Samp$Flag_LAB[FM_Samp$event_id == "SHR_2021-05-10 09:55:00"] <- 1
FM_Samp$Comment_SAMP[FM_Samp$event_id == "SHR_2021-05-10 09:55:00"] <- "LAB"

FM_Samp$Flag_LAB[FM_Samp$event_id == "STTD_2021-05-11 09:31:00"] <- 1
FM_Samp$Comment_LAB[FM_Samp$event_id == "STTD_2021-05-11 09:31:00"] <- "LAB"

FM_Samp$Flag_LAB[FM_Samp$event_id == "SHR_2021-05-24 09:44:00"] <- 1
FM_Samp$Comment_LAB[FM_Samp$event_id == "SHR_2021-05-24 09:44:00"] <- "LAB"

FM_Samp$Flag_LAB[FM_Samp$event_id == "SHR_2022-01-25 09:28:00"] <- 1
FM_Samp$Comment_LAB[FM_Samp$event_id == "SHR_2022-01-25 09:28:00"] <- "LAB"

FM_Samp$Flag_SAMP[FM_Samp$event_id == "SHR_2024-06-18 08:46:00"] <- 3
FM_Samp$Comment_SAMP[FM_Samp$event_id == "SHR_2024-06-18 08:46:00"] <- "FM"
#very high number of revs for flowmeter - verified on data sheet


```

17. Make a table of summarized flowmeter values for different conditions
    * Use Flowdiff/Settime to look for outliers
    * Use inundation as a grouping variable for STTD but not SHR
    
STTD
```{r message = FALSE, warning = FALSE}
### Table of values
# Calculate median, lower and upper ranges for each day
Flow.sum.STTD <- samp3 %>%
  left_join(inundation4) %>%
  left_join(wy)%>%
  filter(!is.na(Flowdiff), !is.na(Inundation2), Station=="STTD") %>%
    mutate(FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Regular", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Reg", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "REG", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Low", "LOW")) %>% 
  mutate(Flow_s = Flowdiff/SetTime) %>%
  group_by(Station, FlowMeterSpeed, WYClass, Inundation2) %>%
  summarize(n= n(),
            min.Flowdiff = min(Flowdiff),
            max.Flowdiff = max(Flowdiff),
            median.Flowdiff = median(Flowdiff),
            min.Flowdiff_s = min(Flow_s),
            max.Flowdiff_s = max(Flow_s),
            median.Flowdiff_s = median(Flow_s),
            Flow_Q1 = quantile(Flowdiff, probs = 0.25),
            Flow_Q3 = quantile(Flowdiff, probs = 0.75),
            Flow_UL = Flow_Q3 + 1.5 * (Flow_Q3-Flow_Q1),
            Flow_s_Q1 = quantile(Flow_s, probs = 0.25), 
            Flow_s_Q3 = quantile(Flow_s, probs = 0.75),
            Flow_s_UL = Flow_s_Q3 + 1.5 * (Flow_s_Q3-Flow_s_Q1),
            Flow_s_MAD = mad(Flow_s))


Flow.sum.STTD %>%
  kbl() %>%
  kable_styling()

```

Sherwood
```{r message = FALSE, warning = FALSE}
Flow.sum.SHR <- samp3 %>%
  left_join(inundation4) %>%
  left_join(wy)%>%
      mutate(FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Regular", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Reg", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "REG", "REGULAR"),
         FlowMeterSpeed = replace(FlowMeterSpeed, FlowMeterSpeed == "Low", "LOW")) %>% 
  mutate(Flow_s = Flowdiff/SetTime) %>%
  group_by(Station, FlowMeterSpeed, WYClass) %>%
  filter(!is.na(Flowdiff), Station == "SHR") %>%
  summarize(n= n(),
            min.Flowdiff = min(Flowdiff),
            max.Flowdiff = max(Flowdiff),
            median.Flowdiff = median(Flowdiff),
            min.Flowdiff_s = min(Flow_s),
            max.Flowdiff_s = max(Flow_s),
            median.Flowdiff_s = median(Flow_s),
            Flow_Q1 = quantile(Flowdiff, probs = 0.25),
            Flow_Q3 = quantile(Flowdiff, probs = 0.75),
            Flow_UL = Flow_Q3 + 1.5 * (Flow_Q3-Flow_Q1),
            Flow_s_Q1 = quantile(Flow_s, probs = 0.25), 
            Flow_s_Q3 = quantile(Flow_s, probs = 0.75),
            Flow_s_UL = Flow_s_Q3 + 1.5 * (Flow_s_Q3-Flow_s_Q1),
            Flow_s_MAD = mad(Flow_s))

Flow.sum.SHR %>%
    kbl() %>%
  kable_styling()

```

18. Apply outlier definitions to data

* Merge table (above) with rest of the data so averages are based on each individual collected sample and not multiplied for each organism in a sample. 
* Calculate modified Z score, then label each row as outlier = Tukey, MAD, or BOTH.
* Do this separately for STTD and SHR
```{r}
Flow.outlier.STTD <- left_join(FM_Samp, Flow.sum.STTD) %>%
    mutate(Flow_s = Flowdiff/SetTime) %>%
  filter(!is.na(Flowdiff), !is.na(Inundation2), Station=="STTD") %>%
  group_by(Station, FlowMeterSpeed, WYClass) %>%
  mutate(Flow_modZ = abs(0.6745*(Flow_s - median.Flowdiff_s)/Flow_s_MAD),
         Flow_Outlier = ifelse((Flow_s > Flow_UL) & Flow_modZ > 3.5, "Both", ifelse(Flow_s > Flow_UL, "Tukey", ifelse(Flow_modZ > 3.5, "MAD", 
            "None"))))

summary(factor(Flow.outlier.STTD$Flow_Outlier))

Flow.outlier.SHR <- left_join(FM_Samp, Flow.sum.SHR) %>%
  mutate(Flow_s = Flowdiff/SetTime) %>%
  filter(!is.na(Flowdiff),  Station=="SHR") %>%
  group_by(Station, FlowMeterSpeed, WYClass) %>%
  mutate(Flow_modZ = abs(0.6745*(Flow_s - median.Flowdiff_s)/Flow_s_MAD),
         Flow_Outlier = ifelse((Flow_s > Flow_UL) & Flow_modZ > 3.5, "Both", ifelse(Flow_s > Flow_UL, "Tukey", ifelse(Flow_modZ > 3.5, "MAD", 
            "None"))))

summary(factor(Flow.outlier.SHR$Flow_Outlier))
```

19. Plot outliers - STTD
```{r}
FlowBoxMonth1a <- ggplot(Flow.outlier.STTD) + geom_boxplot(aes(x = WYClass, y = Flow_s)) +  scale_fill_viridis(discrete = TRUE) + theme_bw() 
ggplotly(FlowBoxMonth1a)


outlierWY1a <- ggplot(Flow.outlier.STTD, aes(x = WYClass, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) +
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))


outlierMonth1a <- ggplot(Flow.outlier.STTD, aes(x = Month, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(FlowBoxMonth1a, outlierWY1a, outlierMonth1a)
```

Sherwood outliers
```{r}
FlowBoxMonth1b <- ggplot(Flow.outlier.SHR) + geom_boxplot(aes(x = WYClass, y = Flow_s)) +  scale_fill_viridis(discrete = TRUE) + theme_bw() 
ggplotly(FlowBoxMonth1b)


outlierWY1b <- ggplot(Flow.outlier.SHR, aes(x = WYClass, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) +   
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))


outlierMonth1b <- ggplot(Flow.outlier.SHR, aes(x = Month, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) +   
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(FlowBoxMonth1b, outlierWY1b, outlierMonth1b)
```

Merge two files
```{r}
Flow.outlier <- rbind(Flow.outlier.SHR, Flow.outlier.STTD)
```


20. Flowmeter Flags and New Flowdiff calculated

Cutoffs based on looking at plots and Upper limits calculated above. 

Flag   | Description                | Outlier                   | Flowdiff
Flag 3 | Highly Suspect             | Both                      | < 200 or >= 70000 or >= 50000 if inundation = FALSE
Flag 2 | Suspect                    | Tukey or MAD but not Both | 200-999 
Flag 1 | Acceptable                 |                           | 
No flag| Acceptable                 | None                      | 1000-59999

21. Add comment about what was flagged (FM for flowmeter-related)
22. Calculate replacement values where Flag_FM = 3 or Comment_SAMP = FM
  * Replacement values are median standardized flowdiff * settime based on groupings of Station, Flowmeter Speed, Water Year Class

```{r Flag-Replace-Flowmeters}
################## Flag Flowmeter data: Flag_QC1 ############################
# Mutate QC Flags 
# Conditions for Flag 3: Outlier = Both | Flowdiff < 200 | Flowdiff >= 50000 when no inundation | Flowdiff >= 70000
# Conditions for Flag 2: Outlier = Tukey | Outlier = MAD (but not Both) |Flowdiff between 200-999 
# Conditions for Flag 1: Acceptable data

# If Flag 3, calculate FlowdiffAdj = median.Flowdiff_s * SetTime

FlowmeterQAQC <- Flow.outlier %>%
  filter(!is.na(Flowdiff), Flow_modZ>=0.00000001) %>%
  mutate(Flag_FM = ifelse(Flow_Outlier=="Both" | Flowdiff < 200 | (Flowdiff >= 50000 & Inundation2 ==FALSE) | Flowdiff >= 70000 , 3, 
                          ifelse(Flow_Outlier %in% c("Tukey", "MAD") | Flowdiff < 1000, 2,
                                 "0")),
         Comment_FM = ifelse(Flag_FM %in% c(2,3), "FM", "0"),
         FlowdiffAdj = ifelse(Flag_FM == 3 | Comment_SAMP == "FM", median.Flowdiff_s * SetTime, Flowdiff))

```

Summarize number of outliers flagged

```{r}
FlowmeterQAQC$Flag_FM <- ordered(FlowmeterQAQC$Flag_FM)
print("Flowmeter Outliers")
summary(factor(FlowmeterQAQC$Flag_FM))
print("Sampling Outliers")
summary(factor(FlowmeterQAQC$Flag_SAMP))
print("Lab Outliers")
summary(factor(FlowmeterQAQC$Flag_LAB))
```
23. Plot changes and flags

```{r}
outlierPreFlow <- ggplot(FlowmeterQAQC, aes(x = factor(Month), y = Flowdiff,  text = paste('Date: ', Date,'<br>Inundation: ', Inundation2, '<br>Set time: ', SetTime), color = factor(Flag_FM))) + geom_point(size = 2.5, alpha = 0.5) +  scale_color_viridis(discrete = TRUE) + labs(title = "Flowmeter Pre-QC") +
  facet_wrap(~Station) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlierPreFlow)

outlierPostFlow <- ggplot(FlowmeterQAQC, aes(x = factor(Month), y = FlowdiffAdj, color = factor(Flag_FM))) + geom_point(size = 2.5, alpha = 0.5) + scale_color_viridis(discrete = TRUE)  + labs(title = "Flowmeter Post-QC") +
  facet_wrap(~Station) +   
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(outlierPreFlow, outlierPostFlow, nrow = 2)
```

24. Plot before and after for SAMP
```{r}
outlierPreSamp <- ggplot(Flow.outlier, aes(x = factor(Month), y = Flowdiff, color = factor(Flag_SAMP))) + geom_point(size = 3, alpha = 0.8) + scale_color_viridis(discrete = TRUE) + labs(title = "Sampling Pre-QC") +
  facet_wrap(~Station) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

outlierPostSamp <- ggplot(FlowmeterQAQC, aes(x = factor(Month), y = FlowdiffAdj, color = factor(Flag_SAMP))) + geom_point(size = 3, alpha = 0.8) + scale_color_viridis(discrete = TRUE) + labs(title = "Sampling Post-QC") +
  facet_wrap(~Station) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(outlierPreSamp, outlierPostSamp, nrow = 2)
```

## QAQC: CPUE

25. Calculate Vol and CPUE, Vol_adj, CPUEAdj

```{r}
NetArea = 0.4572 * 0.25
FlowQAQC_CPUE <- FlowmeterQAQC %>%
  mutate(Volume = ifelse(FlowMeterSpeed == "REGULAR",
                         Flowdiff * 26873 * NetArea/999999,
                         ifelse(FlowMeterSpeed == "LOW",
                                Flowdiff * 57560 * NetArea/999999, NA))) %>%
  filter(!is.na(Flowdiff)) %>%
  mutate(CPUE= round(Count/Volume,3))%>%
    
  mutate(VolumeAdj = ifelse(FlowMeterSpeed == "REGULAR",
                         FlowdiffAdj * 26873 * NetArea/999999,
                         ifelse(FlowMeterSpeed == "LOW",
                                FlowdiffAdj * 57560 * NetArea/999999, NA))) %>%
  mutate(CPUEAdj = round(Count/VolumeAdj,3))%>%
  arrange(Datetime)

```

26. Look at distribution of CPUE values
```{r}
FlowQAQC_CPUE$Month <- factor(FlowQAQC_CPUE$Month)
FlowQAQC_low <- filter(FlowQAQC_CPUE, CPUE < 10)
FlowQAQC_high <- filter(FlowQAQC_CPUE, CPUE > 10)

# Histogram of values
CPUEHist <- ggplot(FlowQAQC_CPUE, aes(CPUEAdj)) + geom_histogram(binwidth = 1) +
  facet_wrap(~Station) + theme_bw()

# SHR vs STTD boxplot
CPUEBox <- ggplot(FlowQAQC_CPUE) + geom_boxplot(aes(x = Station, y = CPUEAdj)) + theme_bw()

# SHR vs STTD boxplot by month
CPUEBoxMonth <- ggplot(FlowQAQC_CPUE) + geom_boxplot(aes(x = Month, y = CPUEAdj, color = Month)) + facet_wrap(~Station) + scale_color_viridis(discrete = TRUE) + theme_bw() 

grid.arrange(CPUEHist, CPUEBox)

# SHR vs STTD boxplot by year
CPUEBoxYear <- ggplot(FlowQAQC_CPUE) + geom_boxplot(aes(x = factor(WY), y = CPUEAdj)) + facet_grid(Station~.) + theme_bw() 
ggplotly(CPUEBoxYear)

grid.arrange(CPUEBoxMonth, CPUEBoxYear)
```

CPUE and Count Plots
```{r fig.width = 10, fig.height = 8}
# Species
inv_summary <-FlowQAQC_CPUE %>%
  filter(WY>1998) %>%
  group_by(Station, WYClass, TaxonName) %>%
  summarize(median.CPUE = median(CPUEAdj),
            max.CPUE = max(CPUEAdj),
            max.Count = max(Count),
            median.Count = median(Count),
            total.CPUE = sum(CPUEAdj),
            n = n()) %>%
  filter(n > 2)

# All CPUE
speciesCPUE <- ggplot(FlowQAQC_CPUE, aes(TaxonName, CPUEAdj)) + 
  geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
  scale_color_viridis(discrete=TRUE) +
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15)) 

ggplotly(speciesCPUE)

# Median CPUE
species2 <- ggplot(inv_summary, aes(TaxonName, median.CPUE)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

# MaxCPUE
species3 <- ggplot(inv_summary, aes(TaxonName, max.CPUE)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))


# Median Count
medCount <- ggplot(inv_summary, aes(TaxonName, median.Count)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

# Max Count
maxCount <- ggplot(inv_summary, aes(TaxonName, max.Count)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(species2, species3)
grid.arrange(medCount, maxCount)

# Which are the outliers?
ggplotly(maxCount)

```


27. Check on and change any known errors
```{r}
checkCounts <- filter(FlowQAQC_CPUE, Count>1000)
# Match up with datasheet 

checkCPUE <- filter(FlowQAQC_CPUE, CPUEAdj > 100)
checkCPUE2 <- filter(FlowQAQC_CPUE, is.na(CPUEAdj))
```

28. Look at summary table of adjusted CPUE values
```{r, eval = FALSE}
### Table of values
CPUE.adj.table <- FlowQAQC_CPUE %>%
  group_by(Station, FlowMeterSpeed, WYClass, TaxonName) %>%
  summarize(min.CPUE = min(CPUEAdj),
            median.CPUE = median(CPUEAdj),
            max.CPUE = max(CPUEAdj),
            n = n())

CPUE.adj.table %>%
    kbl() %>%
  kable_styling()

```

29. Determine and apply flags and cutoffs
    * In this case, used plots to decide to flag CPUE values > 70 

30. Add comment about what was flagged (Comment_CPUE = CPUE)
```{r}
CPUEQAQC <- FlowQAQC_CPUE %>%
  mutate(Flag_CPUE = ifelse(CPUEAdj>70, 2, "0" ),
         Comment_CPUE = ifelse(CPUEAdj>70, "CPUE", "0"))
```

31. Plot outliers
```{r}
outlier2a <- ggplot(CPUEQAQC, aes(x = Month, y = CPUE, color = Flag_FM)) + geom_point(size = 2.5, alpha = 0.5) + scale_color_viridis(discrete = "TRUE") + facet_wrap(~Station) + labs(title = "CPUE calculated with original Flowmeter values") + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

outlier2b <- ggplot(CPUEQAQC, aes(x = Month, y = CPUEAdj, color = Flag_CPUE)) + geom_point(size = 2.5, alpha = 0.5) + scale_color_viridis(discrete = "TRUE") + facet_wrap(~Station) +   labs(title = "CPUE calculated with adjusted Flowmeter values + CPUE flags") + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlier2b)

grid.arrange(outlier2a, outlier2b)

```

## Clean up display of dataset
32. Join taxa and clean up order of variables
```{r}
CPUEQAQC <- FlowQAQC_CPUE %>%
  mutate(Flag_CPUE = ifelse(CPUEAdj>70, 2, "0" ),
         Comment_CPUE = ifelse(CPUEAdj>70, "CPUE", "0"))
```


```{r}
# Rename some columns, remove unnecessary columns
inv_select <- CPUEQAQC %>%
  dplyr::rename(Inundation = Inundation2,
         FlagPhys = Flag_PQC,
         FlagSamp = Flag_SAMP,
         FlagLab = Flag_LAB,
         FlagFM = Flag_FM, 
         FlagCPUE = Flag_CPUE,
         CommentPhys = Comment_PQC,
         CommentSamp = Comment_SAMP,
         CommentLab = Comment_LAB,
         CommentFM = Comment_FM,
         CommentCPUE = Comment_CPUE) %>%
  select(c(event_id, Datetime, Station,
           WY, WYClass, Inundation,
           WeatherCode, Tide, MicrocystisVisualRank,
           WaterTemperature:Turbidity,
           ConditionCode,FieldComments, 
           SetTime, FlowMeterSpeed, FlowMeterStart, FlowMeterEnd, Flowdiff, FlowdiffAdj, Volume, VolumeAdj,LabComments, Category, TaxonName, LifeStage, Count, CPUE, CPUEAdj, FlagPhys, CommentPhys, FlagSamp, CommentSamp, FlagLab, CommentLab, FlagFM, CommentFM, FlagCPUE, CommentCPUE)) 

# Modify some taxa to join correctly
inv_select$TaxonName[inv_select$TaxonName == "affinis"]  <- "Gambusia affinis" 
inv_select$TaxonName[inv_select$TaxonName == "Hydrachnidae"]  <- "Hydrachnidia" 
inv_select$LifeStage[inv_select$TaxonName == "Actinopterygii - eggs"]  <- "eggs"
inv_select$TaxonName[inv_select$TaxonName == "Actinopterygii - eggs"]  <- "Actinopterygii" 

############## CHECK
# Join taxa table to bring organism ID back in
inv_all <- left_join(inv_select, tax) %>%
  filter(!is.na(TaxonName))

# Rearrange 
inv_arrange <- inv_all %>% 
  select(c(event_id:Turbidity, SetTime:VolumeAdj, ConditionCode:FieldComments, LabComments:Category, OrganismID, TaxonRank, TaxonName, CommonName, LifeStage:CommentCPUE))
```


33. Merge flags and comments into one flag, one comment column, replace "" with NA
```{r}
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, "0")
}

inv_final <- inv_arrange %>%
  mutate(LifeStage = replace(LifeStage, LifeStage == "N/A", NA)) %>%
  mutate(across(c(FlagSamp:CommentCPUE), list(empty_as_na)) ) %>%
  select(-c(FlagSamp:CommentCPUE)) %>%
  rename(FlagSamp = FlagSamp_1, 
         FlagLab = FlagLab_1,
         FlagFM = FlagFM_1,
         FlagCPUE = FlagCPUE_1,
         CommentSamp = CommentSamp_1, 
         CommentLab = CommentLab_1,
         CommentFM = CommentFM_1,
         CommentCPUE = CommentCPUE_1) %>%
  mutate(QCFlags = paste(FlagSamp, " ", FlagLab, " ", FlagFM, " ", FlagCPUE)) %>%
  mutate(QCComments = paste(CommentSamp, " ", CommentLab, " ",CommentFM, " ",  CommentCPUE)) %>%
  relocate(FlowMeterSpeed, .before = "FlowMeterStart")

## Filter rows with a 7
threes <- inv_final%>%
  filter(grepl("3", QCFlags ))
```


34. Calculate proportion of data flagged
```{r}
## Amount replaced
print(paste0(round(nrow(threes)/nrow(inv_final)*100,3), "% highly suspicious"))
```

35. Write file
```{r}
# Write File
# quote = TRUE important for EML writing
today = format(today(),"%Y%m%d")
write.csv(inv_final, paste0("R_write/Drift/Drift_data_", today, ".csv"), row.names = FALSE, quote = TRUE)
```

36. Write R-markdown file
