---
title: "Ichthyo_QAQC_update"
author: "Lisa Vance"
date: "2026-05-18"
output: html_document
---

YBFMP Ichthyoplankton Data 1998-2024

```{r set-options, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Set up and load libraries

```{r setup, results=FALSE, warning = FALSE, message = FALSE}
rm(list=ls(all=TRUE))

library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(gridExtra)
library(stringr)
library(plotly)
library(viridis)
library(kableExtra)
library(lubridate)
library(readxl)
library(readr)
library(anytime)
library(magrittr)
library(janitor)
library(tidylog)
```

2. Load data and check variable types
Note data originally came from both MS Access and MS Excel
Physical data is shared between drift and ichthyo since tows are done at the same time and site

```{r load, results = FALSE, warning = FALSE, message = FALSE}

################# bring in qa/qc physical data #################################
PhysData <-read_csv("R_write/LT_phys_qc_20260512.csv")

############# load previously published QA/QC'd data to eventually bind with new data #############
prevpub <- read_csv("R_write/Ichthyo/Ichthyo_data_20241209.csv")
full <- read_csv("R_write/IchFullData1998-2022.csv")
noted <- read_csv("R_write/IchSamplingQAQC_ Notated_2024-10-18.csv")

###### load newest sample and catch data #########
samp <- read_csv("ichthyo data/IchSampDataExcel_20260429.csv", skip =1)
lab23 <- read_csv("ichthyo data/IchLabExcelData_20260429.csv", skip =1)
lab24 <- read_csv("ichthyo data/Ich_20240108-20240617.csv")



########### species codes and crosswalk ###########
IEPFish <- read_csv("ichthyo data/IEP FISH CODE.csv")
Species <- read_csv("ichthyo data/TblLarvalLookUpV2.csv")

######## Bring in Inundation data and Water Year type Data ###################
wy <- read_csv("WaterYearType_CDEC.csv") 
inundation <- read_csv("Yolo_Bypass_Inundation_1998-2024.csv")

```


3.  Rename some columns and drop any unnecessary columns
```{r filter, results = FALSE, message = FALSE, warning = FALSE}

Species <- Species %>%
  rename (SpeciesCode = "Code",
          CommonName = "Species",
          ScientificName = "Scientific Name")

IEPFish <- IEPFish %>%
  select(-c("...5":"...17"))

lab23 <- lab23 %>% 
    rename(Program = 'Measuring program short name',
         Date = 'Sampling Event Date',
         Time = 'Sampling Event Time',
         Station = 'Sampling Area Number',
         SamplingNumber = 'Sampling Event Number',
         SampleID = 'Sample ID',
         OrganismType = 'Observation Type Short Name',
         OrganismGroup = 'Attribute',
         ScientificName = 'Observable',
         TotalCountSpecies = 'Value...10',
         TL = 'Value...11',
         FL = 'Value...12',
         LifeStage = 'Value...13',
         LarvalLifeStage = '...14',) %>%
  filter(!is.na(Program)) %>%
  select(-c(SamplingNumber, SampleID, Program, OrganismType, OrganismGroup))

lab24 <- lab24 %>% 
    select(-c(measuring_program, sam_code, short_name)) %>% 
  rename(Date = date,
         Time = time,
         Station = station,
         SampleID = sample_id,
         Attribute = attribute,
         ScientificName = observable,
         TotalCountSpecies = count,
         LifeStageAbb = life,
         LarvalLifeStage = larval_life_stage)

PhysData <- PhysData %>%
  select(-c(MeterSetTime, FlowMeterStart, FlowMeterEnd, FlowMeterSpeed, ConditionCode))
#these flow meter values relate to drift data, not ichthyo

samp2<- samp %>% 
    rename(Program = 'Measuring program short name',
         Date = 'Sampling Event Date',
         Time = 'Sampling Event Time',
         Station = 'Sampling Area Number',
         SamplingNumber = 'Sampling Event Number',
         ConditionCode = 'Condition Code',
         SamplingAltered = 'Sampling Altered',
         MeterSetTime = 'Set Time',
         FlowMeterStart = 'Flow Meter Start',
         FlowMeterEnd = 'Flow Meter End',
         FlowMeterSpeed = 'Flow Meter Speed',
         EnteredBy = 'Entered by',
         QAQCBy = "QAQC'd by",
         FieldCommentsExcel = 'Field Comments',
         LabComments = 'Lab Comments',
         MeshSize = 'Observation Area Name',
         Observation = 'Observation Area Number',
         SpotNumber = 'Spot Number',
         SpotCode = 'Spot Code (original/duplicate)') %>% 
  select(-c(Program, SpotCode, SpotNumber, EnteredBy, QAQCBy, SamplingNumber)) %>% 
  add_column(TowLocation = "Center", .after = "FlowMeterSpeed")
#adding a tow location column to match prior access data since there were some special studies where the location varied

SpeciesUpdate <- Species %>%
  add_row(SpeciesCode = "UNSS" , CommonName = "silverside", ScientificName = "Menidia sp.") %>%
  add_row(SpeciesCode = "POM", CommonName = "Unid Crappie", ScientificName = "Pomoxis sp.") %>%
  add_row(SpeciesCode = "RAIKIL", CommonName = "Rainwater Killifish", ScientificName = "Lucania parva") %>%
  add_row(SpeciesCode = "KLF", CommonName = "Killifish", ScientificName = "Cyprinodontidae spp.")

# removing TSE - threadfin shad eggs and ASE american shad eggs in this version 
# since it does not appear contractor has made a comment re: eggs
# also the "animal tissue, baby clam, etc" is removed since those NA don't match
# correctly when tied to the excel data
SpeciesUpdate2 <- SpeciesUpdate %>%
  filter(!is.na(ScientificName)) %>%
  filter(SpeciesCode != "TSE") %>%
  filter(SpeciesCode != "ASE") %>% 
  filter(SpeciesCode != "CHNF") %>% 
  filter(SpeciesCode != "CHNA") %>%
  filter(SpeciesCode != "CHNL") %>%
  filter(SpeciesCode != "CHNM") %>%
  filter(SpeciesCode != "CHNS") %>%
  filter(SpeciesCode != "CHNT") %>%
  filter(SpeciesCode != "CHNW")


#look at structure of each df before joining/binding
str(full)
str(lab23)
str(lab24)
str(PhysData)
str(samp2)
str(SpeciesUpdate2)


```


```{r}
lab23$Date<-as.Date(lab23$Date,"%m/%d/%Y")
lab23$Year <- year(lab23$Date)
lab23$Month <- month(lab23$Date)
mymonths <- c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec")
lab23$MonthAbb <- mymonths[ lab23$Month ]
lab23$Datetime = paste(lab23$Date, lab23$Time)
lab23$Datetime <- ymd_hm(lab23$Datetime)
lab23$Time <- strptime(lab23$Time, format = "%H:%M", tz = "") %>%
  strftime(lab23$Time, format = "%H:%M:%S", tz = "", usetz = FALSE,
           digits = getOption("digits.secs"))
lab23$Time <- hms::as_hms(lab23$Time)

lab24$Date<-as.Date(lab24$Date,"%m/%d/%Y")
lab24$Year <- year(lab24$Date)
lab24$Month <- month(lab24$Date)
lab24$MonthAbb <- mymonths[ lab24$Month ]
lab24$Datetime = paste(lab24$Date, lab24$Time)
lab24$Datetime <- ymd_hms(lab24$Datetime)
lab24$Time <- strptime(lab24$Time, format = "%H:%M", tz = "") %>%
  strftime(lab24$Time, format = "%H:%M:%S", tz = "", usetz = FALSE,
           digits = getOption("digits.secs"))
lab24$Time <- hms::as_hms(lab24$Time)

lab24ed <- lab24 %>% 
  rename(TL = total_length,
         FL = fork_length,
         LifeStage = LifeStageAbb) %>% 
  mutate(FL = as.numeric(FL),
         TL = as.numeric(TL),
         TotalCountSpecies = as.numeric(TotalCountSpecies)) %>% 
  select(-(SampleID))

lab23ed <- lab23 %>% 
    mutate(FL = as.numeric(FL),
         TL = as.numeric(TL),
         TotalCountSpecies = as.numeric(TotalCountSpecies))

str(lab23)
str(lab24ed)

labdata <- bind_rows(lab23ed, lab24ed) %>% 
  filter(year(Date) > 2022)

labdata$ScientificName <- str_replace_all(labdata$ScientificName, "Menidia berylina", "Menidia beryllina")

samp2$Date<-as.Date(samp2$Date,"%m/%d/%Y")
samp2$Year <- year(samp2$Date)
samp2$Month <- month(samp2$Date)
samp2$MonthAbb <- mymonths[ samp2$Month ]
samp2$Datetime = paste(samp2$Date, samp2$Time)
samp2$Datetime <- ymd_hms(samp2$Datetime)
samp2$Time <- strptime(samp2$Time, format = "%H:%M", tz = "") %>%
  strftime(samp2$Time, format = "%H:%M:%S", tz = "", usetz = FALSE,
           digits = getOption("digits.secs"))
samp2$Time <- hms::as_hms(samp2$Time)

sampcatch <- left_join(labdata, samp2) %>% 
  mutate(event_id = paste0(Station, "_", Datetime)) %>% 
  relocate(event_id)

phys <- PhysData %>% 
  filter(Station == "STTD") %>% 
  filter(MonthAbb != "Jul") %>% 
  filter(MonthAbb != "Aug") %>% 
  filter(MonthAbb != "Sep") %>%
  filter(MonthAbb != "Oct") %>% 
  filter(MonthAbb != "Nov") %>% 
  filter(MonthAbb != "Dec")
  


sampcatchphys <- left_join(phys, sampcatch)

sampcatchspecies <- left_join(sampcatchphys, SpeciesUpdate2)

fulldata <- sampcatchspecies %>% 
   relocate(event_id, Datetime, Date, Time, Station, YSI, WeatherCode, Tide, FlowDirection,
           ConditionCode, VegetationRank, MicrocystisVisualRank, WaterTemperature, Secchi, Conductivity, SpCnd, pH, DO, Turbidity, MeterSetTime, FlowMeterSpeed, FlowMeterStart, FlowMeterEnd, TowLocation)


```

Narrow down to unique samples
```{r}
sampUnique <- fulldata %>%
  unique() %>%
  arrange(Datetime)

samp_catch_physMerge <- fulldata %>%
  mutate(WY = ifelse(Month >9, Year + 1, Year)) %>%
  left_join(wy, by = "WY") %>%
  select(-c(Index, WYType)) %>%
  mutate(Flowdiff = abs(FlowMeterEnd-FlowMeterStart))%>%
  select(-c(TotalCountSpecies)) %>%
  arrange(Datetime)
```

## QAQC 

* QC Flags: 0 = no issues,no comments, blank/1 = Comment but deemed Acceptable/ 2 = Suspect/ 3 = Highly Suspect
* Flag_SAMP: Sampling issues
* Flag_LAB: Lab issues
* Flag_FM: Flowmeter
* Flag_CPUE: CPUE

### QAQC Sampling and Lab issues

12. Export sampling table
13. Look at field sampling comments
14. Flag comments reflecting poor data quality (Flag_SAMP)
15. Add comment about what was flagged
    * FM if issue with flowmeter
    * FMMISSING if lacking flowmeter values
    * SAMP for other sampling issues 
    * Leave blank if no issues
16. Look at lab comments and flag if poor data quality
17. Add Comment_LAB
    * LAB if lab issues (e.g. unable to meet meso or microtally)
    * Leave blank if no issues

* This part is manual. Think about how to make this codeable, change our condition code flags or add a flag as part of data entry? 
* The following section creates blank columns before creating the csv to manually filter through comments and add appropriate notation before reading the file back in at step 18

```{r}

#find duplicated, ensure only looking at each individual event id but including comment extra to ensure no important comment is missed
SamplingQAQC <- fulldata[duplicated(fulldata$event_id) == FALSE,c("event_id", "Datetime","Station","ConditionCode","MeterSetTime","FlowMeterSpeed","FlowMeterStart", "FlowMeterEnd", "TowLocation","SamplingAltered","FieldComments", "LabComments")] %>% 
  arrange(Datetime)


SamplingQAQC$Flag_SAMP <-  ifelse(is.na(SamplingQAQC$FieldComments), 0, "")
SamplingQAQC$Comment_SAMP <-""
SamplingQAQC$Flag_LAB <- ifelse(is.na(SamplingQAQC$LabComments), 0, "")
SamplingQAQC$Comment_LAB <- ""
today <- today()
write.csv(SamplingQAQC, paste("R_write/IchSamplingQAQC_",today, ".csv"))


```

