---
title: "Ichthyo_QAQC_update"
author: "Lisa Vance"
date: "2026-05-18"
output: html_document
---

YBFMP Ichthyoplankton Data 1998-2024

```{r set-options, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Set up and load libraries

```{r setup, results=FALSE, warning = FALSE, message = FALSE}
rm(list=ls(all=TRUE))

library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(gridExtra)
library(stringr)
library(plotly)
library(viridis)
library(kableExtra)
library(lubridate)
library(readxl)
library(readr)
library(anytime)
library(magrittr)
library(janitor)
library(tidylog)
```

2. Load data and check variable types
Note data originally came from both MS Access and MS Excel
Physical data is shared between drift and ichthyo since tows are done at the same time and site

```{r load, results = FALSE, warning = FALSE, message = FALSE}

################# bring in qa/qc physical data #################################
PhysData <-read_csv("R_write/LT_phys_qc_20260512.csv")
Phys2 <- read_csv("drift data/LT_phys_qc_20240118.csv")

############# load previously published QA/QC'd data to eventually bind with new data #############
prevpub <- read_csv("R_write/Ichthyo/Ichthyo_data_20241209.csv")
full <- read_csv("R_write/IchFullData1998-2022.csv")
noted <- read_csv("R_write/IchSamplingQAQC_ Notated_2024-10-18.csv")

###### load newest sample and catch data #########
samp <- read_csv("ichthyo data/IchSampDataExcel_20260429.csv", skip =1)
labcomments <- read_csv("ichthyo data/Ich_20190422-20240617.csv")
#note there was a transciption error for time sent to contractors on COC for 5/6/2019 - corrected
# in csv (9:56 incorrect, 9:53 correct per datasheet)
SamplingData <- read_csv("ichthyo data/TblLarvalSampAccess.csv") #for data 4/22/2019 to jan 2020

########### species codes and crosswalk ###########
IEPFish <- read_csv("ichthyo data/IEP FISH CODE.csv")
Species <- read_csv("ichthyo data/TblLarvalLookUpV2.csv")

######## Bring in Inundation data and Water Year type Data ###################
wy <- read_csv("WaterYearType_CDEC.csv") 
inundation <- read_csv("Yolo_Bypass_Inundation_1998-2024.csv")

```


3.  Rename some columns and drop any unnecessary columns
```{r filter, results = FALSE, message = FALSE, warning = FALSE}
Species <- Species %>%
  rename (SpeciesCode = "Code",
          CommonName = "Species",
          ScientificName = "Scientific Name")

IEPFish <- IEPFish %>%
  select(-c("...5":"...17"))

labupdate <- labcomments %>% 
    select(-c(measuring_program, sam_code, short_name)) %>% 
  rename(Date = date,
         Time = time,
         Station = station,
         SampleID = sample_id,
         Attribute = attribute,
         ScientificName = taxon,
         TotalCountSpecies = count,
         LifeStage = life_stage,
         LarvalLifeStageAbb = larval_life_stage,
         LabComments = original_taxonomist_notes,
         QCReID = qc_reid,
         QC_taxon = qc_taxon,
         QC_comments = qc_notes) %>% 
  select(-SampleID)


PhysData <- PhysData %>%
  select(-c(MeterSetTime, FlowMeterStart, FlowMeterEnd, FlowMeterSpeed, ConditionCode))
#these flow meter values relate to drift data, not ichthyo

samp2<- samp %>% 
    rename(Program = 'Measuring program short name',
         Date = 'Sampling Event Date',
         Time = 'Sampling Event Time',
         Station = 'Sampling Area Number',
         SamplingNumber = 'Sampling Event Number',
         ConditionCode = 'Condition Code',
         SamplingAltered = 'Sampling Altered',
         MeterSetTime = 'Set Time',
         FlowMeterStart = 'Flow Meter Start',
         FlowMeterEnd = 'Flow Meter End',
         FlowMeterSpeed = 'Flow Meter Speed',
         EnteredBy = 'Entered by',
         QAQCBy = "QAQC'd by",
         FieldCommentsExcel = 'Field Comments',
         LabComments = 'Lab Comments',
         MeshSize = 'Observation Area Name',
         Observation = 'Observation Area Number',
         SpotNumber = 'Spot Number',
         SpotCode = 'Spot Code (original/duplicate)') %>% 
  select(-c(Program, SpotCode, SpotNumber, EnteredBy, QAQCBy, SamplingNumber)) %>% 
  add_column(TowLocation = "Center", .after = "FlowMeterSpeed")
#adding a tow location column to match prior access data since there were some special studies where the location varied

SpeciesUpdate <- Species %>%
  add_row(SpeciesCode = "UNSS" , CommonName = "silverside", ScientificName = "Menidia sp.") %>%
  add_row(SpeciesCode = "POM", CommonName = "Unid Crappie", ScientificName = "Pomoxis sp.") %>%
  add_row(SpeciesCode = "RAIKIL", CommonName = "Rainwater Killifish", ScientificName = "Lucania parva") %>%
  add_row(SpeciesCode = "KLF", CommonName = "Killifish", ScientificName = "Cyprinodontidae spp.")

# removing TSE - threadfin shad eggs and ASE american shad eggs in this version 
# since it does not appear contractor has made a comment re: eggs
# also the "animal tissue, baby clam, etc" is removed since those NA don't match
# correctly when tied to the excel data
SpeciesUpdate2 <- SpeciesUpdate %>%
  filter(!is.na(ScientificName)) %>%
  filter(SpeciesCode != "TSE") %>%
  filter(SpeciesCode != "ASE") %>% 
  filter(SpeciesCode != "CHNF") %>% 
  filter(SpeciesCode != "CHNA") %>%
  filter(SpeciesCode != "CHNL") %>%
  filter(SpeciesCode != "CHNM") %>%
  filter(SpeciesCode != "CHNS") %>%
  filter(SpeciesCode != "CHNT") %>%
  filter(SpeciesCode != "CHNW")


#look at structure of each df before joining/binding
str(full)
str(lab23)
str(lab24)
str(PhysData)
str(samp2)
str(SpeciesUpdate2)


```


```{r}
labupdate$Date<-as.Date(labupdate$Date,"%m/%d/%Y")
labupdate$Year <- year(labupdate$Date)
labupdate$Month <- month(labupdate$Date)
mymonths <- c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec")
labupdate$MonthAbb <- mymonths[ labupdate$Month ]
labupdate$Datetime = paste(labupdate$Date, labupdate$Time)
labupdate$Datetime <- ymd_hms(labupdate$Datetime)
labupdate$Time <- strptime(labupdate$Time, format = "%H:%M", tz = "") %>%
  strftime(labupdate$Time, format = "%H:%M:%S", tz = "", usetz = FALSE,
           digits = getOption("digits.secs"))
labupdate$Time <- hms::as_hms(labupdate$Time)


labupdateed <- labupdate %>% 
  rename(TL = total_length,
         FL = fork_length,
         LifeStageAbb = LarvalLifeStageAbb) %>% 
  mutate(FL = as.numeric(FL),
         TL = as.numeric(TL),
         TotalCountSpecies = as.numeric(TotalCountSpecies))

labupdateed$ScientificName <- str_replace_all(labupdateed$ScientificName, "Menidia berylina", "Menidia beryllina")

samp2$Date<-as.Date(samp2$Date,"%m/%d/%Y")
samp2$Year <- year(samp2$Date)
samp2$Month <- month(samp2$Date)
samp2$MonthAbb <- mymonths[ samp2$Month ]
samp2$Datetime = paste(samp2$Date, samp2$Time)
samp2$Datetime <- ymd_hms(samp2$Datetime)
samp2$Time <- strptime(samp2$Time, format = "%H:%M", tz = "") %>%
  strftime(samp2$Time, format = "%H:%M:%S", tz = "", usetz = FALSE,
           digits = getOption("digits.secs"))
samp2$Time <- hms::as_hms(samp2$Time)

labupdate2020on <- labupdateed %>% 
  filter(year(Date) > 2019)

labupdate2019 <- labupdateed %>% 
  filter(year(Date) == 2019)

phys2 <- Phys2 %>% 
  filter(Date > "2019-04-21" & Date < "2020-01-28") %>% 
  select(-c(MeterSetTime, FlowMeterStart, FlowMeterEnd, FlowMeterSpeed, ConditionCode)) %>% #flowmeter values relate to drift
  filter(Station == "STTD") %>% 
  filter(MonthAbb != "Jul") %>% 
  filter(MonthAbb != "Aug") %>% 
  filter(MonthAbb != "Sep") %>%
  filter(MonthAbb != "Oct") %>% 
  filter(MonthAbb != "Nov") %>% 
  filter(MonthAbb != "Dec")

phys3 <- Phys2 %>% 
  filter(Date > "2019-12-31") %>% 
  select(-c(MeterSetTime, FlowMeterStart, FlowMeterEnd, FlowMeterSpeed, ConditionCode)) %>% #flowmeter values relate to drift
  filter(Station == "STTD") %>% 
  filter(MonthAbb != "Jul") %>% 
  filter(MonthAbb != "Aug") %>% 
  filter(MonthAbb != "Sep") %>%
  filter(MonthAbb != "Oct") %>% 
  filter(MonthAbb != "Nov") %>% 
  filter(MonthAbb != "Dec")

sampphys <- left_join(phys2, SamplingData) %>% 
  select(-c("2ndStartTime_East", "2ndStopTime_East", StartTime, StopTime,"LarvalStartMeter_Mid_West",
            "LarvalEndMeter_Mid_West","LarvalStartMeter_Near_West","LarvalEndMeter_Near_West",
            "LarvalStartMeter_Mid_East","LarvalEndMeter_Mid_East","LarvalStartMeter_Near_East",
            "LarvalEndMeter_Near_East","LarvalStartMeter_Mid_Bottom","LarvalEndMeter_Mid_Bottom",
            Bottom_StartTime, LarvalCode, EnteredBy, "QA/QC'dBy", SampleVolume)) %>% 
  rename(FlowMeterStart = LarvalStartMeter,
         FlowMeterEnd = LarvalEndMeter,
         MeterSetTime = SetTime) %>% 
  add_column(TowLocation = "Center", .after = "FlowMeterSpeed")

#this takes care of the overlap in 2019 between access and excel for phys data and lab data 
sampcatchphys <- left_join(sampphys, labupdate2019) %>% 
  filter(year(Date) == 2019)

#this takes care of 2020 onwards
sampupdate <- samp2 %>% 
  filter(year(Date) < 2025) %>% 
  select(-(LabComments))

#this matches the 2020 onward lab data with sampling data
sampcatch <- left_join(sampupdate, labupdate2020on) %>% 
  mutate(event_id = paste0(Station, "_", Datetime)) %>% 
  relocate(event_id)

#ichthyoplankton tows are not done Jul-Dec, joining the catch creates unnecessary NA if you don't filter
phys <- PhysData %>% 
  filter(Station == "STTD") %>% 
  filter(MonthAbb != "Jul") %>% 
  filter(MonthAbb != "Aug") %>% 
  filter(MonthAbb != "Sep") %>%
  filter(MonthAbb != "Oct") %>% 
  filter(MonthAbb != "Nov") %>% 
  filter(MonthAbb != "Dec")
  

#this only includes 2023 and 2024
sampcatchphys2 <- left_join(phys, sampcatch)

#this is just for 4/22/2019 through end of 2019
sampcatchspecies <- left_join(sampcatchphys, SpeciesUpdate2)

#this is only 2023 and 2024
sampcatchspecies2 <- left_join(sampcatchphys2, SpeciesUpdate2)

#this is for 2020 to 2022
sampcatchphys3 <- left_join(phys3, sampcatch)

sampcatchspecies3 <- left_join(sampcatchphys3, SpeciesUpdate2)

#filter for only jan 2020?
samp2020 <- sampcatch %>% 
  filter(Date == "2020-01-06" | Date == "2020-01-27")

phys2020 <- phys3 %>% 
  filter(Date == "2020-01-06" | Date == "2020-01-27")

jan2020 <- left_join(samp2020, phys2020)

#now bind 2019, 2020 to 2022 and 2023 to 2024 all together for the updated dataset (including lab comments)
allsampcatchspecies <- bind_rows(sampcatchspecies,sampcatchspecies2, sampcatchspecies3)
##find the 2 dates for 2020 and make sure they get added in 

fulldata <- allsampcatchspecies %>% 
   relocate(event_id, Datetime, Date, Time, Station, YSI, WeatherCode, Tide, FlowDirection,
           ConditionCode, VegetationRank, MicrocystisVisualRank, WaterTemperature, Secchi, Conductivity, SpCnd, pH, DO, Turbidity, MeterSetTime, FlowMeterSpeed, FlowMeterStart, FlowMeterEnd, TowLocation) %>% 
  arrange(Datetime)

#for notating qaqc comments later
noted2 <- noted %>% 
  filter(Datetime < "4/22/2019 9:58")
```

Narrow down to unique samples
```{r}
sampUnique <- fulldata %>%
  unique() %>%
  arrange(Datetime)

samp_catch_physMerge <- fulldata %>%
  mutate(WY = ifelse(Month >9, Year + 1, Year)) %>%
  left_join(wy, by = "WY") %>%
  select(-c(Index, WYType)) %>%
  mutate(Flowdiff = abs(FlowMeterEnd-FlowMeterStart))%>%
  select(-c(TotalCountSpecies)) %>%
  arrange(Datetime)
```

## QAQC 

* QC Flags: 0 = no issues,no comments, blank/1 = Comment but deemed Acceptable/ 2 = Suspect/ 3 = Highly Suspect
* Flag_SAMP: Sampling issues
* Flag_LAB: Lab issues
* Flag_FM: Flowmeter
* Flag_CPUE: CPUE

### QAQC Sampling and Lab issues

12. Export sampling table
13. Look at field sampling comments
14. Flag comments reflecting poor data quality (Flag_SAMP)
15. Add comment about what was flagged
    * FM if issue with flowmeter
    * FMMISSING if lacking flowmeter values
    * SAMP for other sampling issues 
    * Leave blank if no issues
16. Look at lab comments and flag if poor data quality
17. Add Comment_LAB
    * LAB if lab issues (e.g. unable to meet meso or microtally)
    * Leave blank if no issues

* This part is manual. Think about how to make this codeable, change our condition code flags or add a flag as part of data entry? 
* The following section creates blank columns before creating the csv to manually filter through comments and add appropriate notation before reading the file back in at step 18

```{r}

#find duplicated, ensure only looking at each individual event id but including comment extra to ensure no important comment is missed
SamplingQAQC <- fulldata[duplicated(fulldata$event_id) == FALSE,c("event_id", "Datetime","Station","ConditionCode","MeterSetTime","FlowMeterSpeed","FlowMeterStart", "FlowMeterEnd", "TowLocation","SamplingAltered","FieldComments", "LabComments")] %>% 
  arrange(Datetime)


SamplingQAQC$Flag_SAMP <-  ifelse(is.na(SamplingQAQC$FieldComments), 0, "")
SamplingQAQC$Comment_SAMP <-""
SamplingQAQC$Flag_LAB <- ifelse(is.na(SamplingQAQC$LabComments), 0, "")
SamplingQAQC$Comment_LAB <- ""
today <- today()
write.csv(SamplingQAQC, paste("R_write/IchSamplingQAQC_",today, ".csv"))


```


Combine new data with previous data for outlier QAQC
```{r}

fullich <- bind_rows(full, fulldata) %>% 
  arrange(Datetime)

#write new csv to include full updated data
write.csv(fullich, "ichthyo data/IchthyoplanktonData 1998-2024.csv")



```

Now that the QAQC file is updated, read that file back in for the next steps
```{r}
SamplingQAQC_fill <- read.csv("R_write/IchSamplingQAQC_.csv")
SamplingQAQC_fill_s <- SamplingQAQC_fill %>%
  select(c(event_id, PhysicalDataID, Flag_SAMP, Comment_SAMP, Flag_LAB, Comment_LAB))
#bring in previous qaqc notated file as well
SamplingQAQC_fill2 <- noted2
SamplingQAQC_fill_s2 <- SamplingQAQC_fill2 %>%
  select(c(event_id, PhysicalDataID, Flag_SAMP, Comment_SAMP, Flag_LAB, Comment_LAB))

SamplingQAQC_fill_sfinal <- bind_rows(SamplingQAQC_fill_s, SamplingQAQC_fill_s2)

```

### QAQC Flowmeters

19. Inundation data

    This data comes from Water Data Library, Stations at Fremont Weir (A02170) and I Street Bridge . If Stage Height > 32ft at Fremont Weir, Inundation = TRUE. If there is no data at Fremont Weir, and Stage Height > 17 ft at I Street Bridge, Inundation = TRUE. There was also 2 days of lag time added in for inundation (variable Inundation2). 
    
* Inundation used to QA/QC STTD data. 

```{r}
#first make sure inundation columns match ich data frame 
inundation <- inundation %>%
  rename(Date = Dates)
inundation$Date<-as.Date(inundation$Date,"%m/%d/%Y")

inundation2 <- inundation %>%
  mutate(Month = month(Date),
         Year = year(Date),
         WY = ifelse(Month > 9, Year + 1, Year)) %>%
  select(-c(Month, Year))

inundation2a <- inundation %>%
  mutate(Month = month(Date),
         Year = year(Date),
         WY = ifelse(Month > 9, Year + 1, Year))

# Modify sample table to include Flowdiff
samp3 <-sampUnique %>%
  mutate(Flowdiff = abs(FlowMeterEnd-FlowMeterStart),
         Flowdiff_s = Flowdiff/MeterSetTime) 

# merged inundation and sampling
inundation_flow <- left_join(samp3, inundation2, by = "Date") %>%
  filter(Flowdiff<100000) %>%
  filter(Station == "STTD") %>%
  mutate(Flowdiff_s = Flowdiff/MeterSetTime,
         Inundation_n = ifelse(Inundation == "TRUE", 40000, 0))

# all inundation
inundation3 <- inundation %>%
  mutate(Inundation_n = ifelse(Inundation == "TRUE", 40000, 0),
         Inundation_n_low = ifelse(Inundation == "TRUE", 5000,0))

# Plot - only inundation events that correspond with sampling date
inplot1 <- ggplot(inundation_flow) + 
  geom_point(aes(x = Date, y = Flowdiff)) + 
  geom_col(aes(x = Date, y = Inundation_n), fill = "blue", linewidth = 2) + 
  labs(title = "Flowdiff with Inundation events corresponding with sample dates") +
  facet_grid(Station~.) +
  theme_bw()

ggplotly(inplot1)

# Plot - All inundation events
inplot2 <- ggplot() +
  geom_point(data = inundation_flow, aes(x = Date, y = Flowdiff)) +
  labs(title = "Flowdiff with all Inundation") +
  geom_col(data = inundation3, aes(x = Date, y = Inundation_n), fill = 
             "blue", alpha = 0.6) + theme_bw()
ggplotly(inplot2)

inplot3 <- ggplot() +
  geom_point(data = inundation_flow, aes(x = Date, y = Flowdiff_s)) +
  labs(title = "Standardized Flowdiff with all Inundation") +
  geom_col(data = inundation3, aes(x = Date, y = Inundation_n_low), fill = 
             "blue", alpha = 0.6) + theme_bw()

grid.arrange(inplot1, inplot2, inplot3, nrow = 3)


# Allow 10 days after last "inundation=TRUE" to also count as inundated. Merge with sample data. 
inundation4 <- inundation2a %>%
  mutate(Inundation2 = ifelse(lag(Inundation, 10) == "TRUE", "TRUE", Inundation)) %>%
  select(c(Date, Month:Inundation2))

inundation4$Inundation2[inundation4$Date=="1998-01-01"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-02"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-03"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-04"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-05"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-06"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-07"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-08"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-09"] <- "FALSE"
inundation4$Inundation2[inundation4$Date=="1998-01-10"] <- "FALSE"

samp_catch_phys <- left_join(samp_catch_physMerge, inundation4)
```

##Flowmeter QA/QC

20. Look at distribution of flowmeter values
```{r}
#make sure case matches for flowmeter speed
samp_catch_phys$FlowMeterSpeed <- str_replace_all(samp_catch_phys$FlowMeterSpeed, "low", "Low")
samp3$FlowMeterSpeed <- str_replace_all(samp3$FlowMeterSpeed, "low", "Low")

# Histogram of values
samp_catch_phys$Month <- ordered(samp_catch_phys$Month)
FlowHist <- ggplot(samp_catch_phys, aes(Flowdiff)) + geom_histogram() +
  facet_wrap(~Station) + theme_bw()
ggplotly(FlowHist)

# SHR vs STTD boxplot
FlowBox <- ggplot(samp_catch_phys) + geom_boxplot(aes(x = Station, y = Flowdiff)) + theme_bw()

# SHR vs STTD boxplot by month
FlowBoxMonth <- ggplot(samp_catch_phys) + geom_boxplot(aes(x = Month, y = Flowdiff, fill = Month)) + facet_wrap(~Station) + scale_fill_viridis(discrete = TRUE) + theme_bw() 

# SHR vs STTD boxplot by year
FlowBoxYear <- ggplot(samp_catch_phys) + geom_boxplot(aes(x = ordered(WY), y = Flowdiff, fill = ordered(WY))) + facet_wrap(~Station) + theme_bw() 

# SetTime vs Flowdiff by station and month
FlowPoint <- ggplot(samp_catch_phys, aes(x = MeterSetTime, y = Flowdiff, color = Month)) + geom_point(size = 3) + facet_wrap(~Station) + theme_bw()

# SetTime vs Flowdiff by station and flowmeter type
FlowPoint2 <- ggplot(samp_catch_phys, aes(x = MeterSetTime, y = Flowdiff, color = FlowMeterSpeed)) + geom_point(size = 3) + facet_wrap(~Station) + theme_bw() + scale_color_viridis(discrete = TRUE) 


grid.arrange(FlowBox, FlowHist)
grid.arrange(FlowBoxMonth, FlowBoxYear)
grid.arrange(FlowPoint, FlowPoint2)
```


21. Make adjustments to known flowmeter errors from looking at any values outside of expected range
```{r}

checkich <- samp_catch_phys %>% filter(Flowdiff > 40000)
checkich2 <- samp_catch_phys %>% filter(Flowdiff <1000)
#checking for later outlier section
checkich3 <- samp_catch_phys %>% filter(Flowdiff <200)

#fix known flowmeter errors
#flowmeter when it reads 0, use as if it reads 1000000

samp_catch_phys$Flowdiff[samp_catch_phys$event_id == "STTD_2015-04-30 08:16:00"] <- 1000000-996009
samp_catch_phys$FlowMeterEnd[samp_catch_phys$event_id == "SHR_2016-03-17 08:25:00"] <- 954818
samp_catch_phys$Flowdiff[samp_catch_phys$event_id == "SHR_2016-03-17 08:25:00"] <- 954818-941900

#comment on data sheet and confirmation with NI, 7/25/2012 STTD flowmeter value was recorded incorrectly, should match flowmeter for zoop
samp_catch_phys$FlowMeterEnd[samp_catch_phys$event_id == "STTD_2012-07-25 10:11:00"] <- 899989
samp_catch_phys$Flowdiff[samp_catch_phys$event_id == "STTD_2012-07-25 10:11:00"] <- 900000-899989

# 1/24/2012 SHR appears to be correct numbers based on datasheet but flow diff very high
# 1/23/2004 SHR appears to be correct numbers based on datasheet but flow diff very high
# 2/16/1999 SHR appears to be correct numbers based on datasheet but flow diff very high

#change the same flowmeter values in samp3
samp3$Flowdiff[samp3$event_id == "STTD_2015-04-30 08:16:00"] <- 1000000-996009
samp3$FlowMeterEnd[samp3$event_id == "SHR_2016-03-17 08:25:00"] <- 954818
samp3$Flowdiff[samp3$event_id == "SHR_2016-03-17 08:25:00"] <- 954818-941900

#comment on data sheet and confirmation with NI, 7/25/2012 STTD flowmeter value was recorded incorrectly, should match flowmeter for zoop
samp3$FlowMeterEnd[samp3$event_id == "STTD_2012-07-25 10:11:00"] <- 899989
samp3$Flowdiff[samp3$event_id == "STTD_2012-07-25 10:11:00"] <- 900000-899989

#remaining outlier values will be flagged in later qa/qc of flowmeter values
```


22. Merge data with lab/sampling QAQC 
  * Replace NA with blank
```{r}

FM_Samp <- left_join(samp_catch_phys, SamplingQAQC_fill_s, by = "event_id") %>%
  mutate(Flag_SAMP = replace(Flag_SAMP, is.na(Flag_SAMP), "" ),
         Comment_SAMP = replace(Comment_SAMP, is.na(Comment_SAMP), ""),
         Flag_LAB = replace(Flag_LAB, is.na(Flag_LAB), ""),
         Comment_LAB = replace(Comment_LAB, is.na(Comment_LAB), "")) %>%
  select(-c(PhysicalDataID.y))%>%
  rename(PhysicalDataID = "PhysicalDataID.x")

```

23. Make a table of summarized flowmeter values for different conditions
    * Use Flowdiff/Settime to look for outliers
    * Use inundation as a grouping variable for STTD but not SHR
    
STTD
```{r}
#removed meter set times that were NA due to errors in code
Flow.sum.STTD <- samp3 %>%
  left_join(inundation4) %>%
  left_join(wy)%>%
  filter(!is.na(Flowdiff), !is.na(Inundation2),!is.na(MeterSetTime), Station=="STTD") %>%
  mutate(Flow_s = Flowdiff/MeterSetTime) %>%
  group_by(Station, FlowMeterSpeed, WYClass, Inundation2) %>%
  summarize(n= n(),
            min.Flowdiff = min(Flowdiff),
            max.Flowdiff = max(Flowdiff),
            median.Flowdiff = median(Flowdiff),
            min.Flowdiff_s = min(Flow_s),
            max.Flowdiff_s = max(Flow_s),
            median.Flowdiff_s = median(Flow_s),
            Flow_Q1 = quantile(Flowdiff, probs = 0.25),
            Flow_Q3 = quantile(Flowdiff, probs = 0.75),
            Flow_UL = Flow_Q3 + 1.5 * (Flow_Q3-Flow_Q1),
            Flow_s_Q1 = quantile(Flow_s, probs = 0.25), 
            Flow_s_Q3 = quantile(Flow_s, probs = 0.75),
            Flow_s_UL = Flow_s_Q3 + 1.5 * (Flow_s_Q3-Flow_s_Q1),
            Flow_s_MAD = mad(Flow_s))

Flow.sum.STTD %>%
  kbl() %>%
  kable_styling()
```


Sherwood (SHR)
```{r}

#removed meter set times that were NA due to errors in code
Flow.sum.SHR <- samp3 %>%
  left_join(inundation4) %>%
  left_join(wy)%>%
  mutate(Flow_s = Flowdiff/MeterSetTime) %>%
  group_by(Station, FlowMeterSpeed, WYClass) %>%
  filter(!is.na(Flowdiff), !is.na(MeterSetTime), Station == "SHR") %>%
  summarize(n= n(),
            min.Flowdiff = min(Flowdiff),
            max.Flowdiff = max(Flowdiff),
            median.Flowdiff = median(Flowdiff),
            min.Flowdiff_s = min(Flow_s),
            max.Flowdiff_s = max(Flow_s),
            median.Flowdiff_s = median(Flow_s),
            Flow_Q1 = quantile(Flowdiff, probs = 0.25),
            Flow_Q3 = quantile(Flowdiff, probs = 0.75),
            Flow_UL = Flow_Q3 + 1.5 * (Flow_Q3-Flow_Q1),
            Flow_s_Q1 = quantile(Flow_s, probs = 0.25), 
            Flow_s_Q3 = quantile(Flow_s, probs = 0.75),
            Flow_s_UL = Flow_s_Q3 + 1.5 * (Flow_s_Q3-Flow_s_Q1),
            Flow_s_MAD = mad(Flow_s))

Flow.sum.SHR %>%
  kbl() %>%
  kable_styling()


```

24.Apply outlier definitions to data

* Merge table (above) with rest of the data so averages are based on each individual collected sample and not multiplied for each organism in a sample. 
* Calculate modified Z score, then label each row as outlier = Tukey, MAD, or BOTH.
* Do this separately for STTD and SHR

```{r}
Flow.outlier.STTD <- left_join(FM_Samp, Flow.sum.STTD) %>%
  mutate(Flow_s = Flowdiff/MeterSetTime) %>%
  filter(!is.na(Flowdiff), !is.na(Inundation2), Station=="STTD") %>%
  group_by(Station, FlowMeterSpeed, WYClass) %>%
  mutate(Flow_modZ = abs(0.6745*(Flow_s - median.Flowdiff_s)/Flow_s_MAD),
         Flow_Outlier = ifelse((Flow_s > Flow_UL) & Flow_modZ > 3.5, "Both", ifelse(Flow_s > Flow_UL, "Tukey",
                                                                                    ifelse(Flow_modZ > 3.5, "MAD",
                                                                                           "None"))))

Flow.outlier.STTD$FlowMeterSpeed <- str_replace_all(Flow.outlier.STTD$FlowMeterSpeed, "low", "Low")

summary(factor(Flow.outlier.STTD$Flow_Outlier))


Flow.outlier.SHR <- left_join(FM_Samp, Flow.sum.SHR) %>%
  mutate(Flow_s = Flowdiff/MeterSetTime) %>%
  filter(!is.na(Flowdiff),  Station=="SHR") %>%
  group_by(Station, FlowMeterSpeed, WYClass) %>%
  mutate(Flow_modZ = abs(0.6745*(Flow_s - median.Flowdiff_s)/Flow_s_MAD),
         Flow_Outlier = ifelse((Flow_s > Flow_UL) & Flow_modZ > 3.5, "Both", ifelse(Flow_s > Flow_UL, "Tukey",
                                                                                    ifelse(Flow_modZ > 3.5,
                                                                                           "MAD","None"))))

summary(factor(Flow.outlier.SHR$Flow_Outlier))
```

25. Plot Outliers - STTD
```{r}
FlowBoxMonth1a <- ggplot(Flow.outlier.STTD) + geom_boxplot(aes(x = WYClass, y = Flow_s)) +  scale_fill_viridis(discrete = TRUE) + theme_bw() 
ggplotly(FlowBoxMonth1a)


outlierWY1a <- ggplot(Flow.outlier.STTD, aes(x = WYClass, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) +
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))


outlierMonth1a <- ggplot(Flow.outlier.STTD, aes(x = Month, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(FlowBoxMonth1a, outlierWY1a, outlierMonth1a)
```

Plot Outliers - Sherwood
```{r}
FlowBoxMonth1b <- ggplot(Flow.outlier.SHR) + geom_boxplot(aes(x = WYClass, y = Flow_s)) +  scale_fill_viridis(discrete = TRUE) + theme_bw() 
ggplotly(FlowBoxMonth1b)


outlierWY1b <- ggplot(Flow.outlier.SHR, aes(x = WYClass, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) +   
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))


outlierMonth1b <- ggplot(Flow.outlier.SHR, aes(x = Month, y = Flow_s, color = Flow_Outlier)) + geom_point(size = 2.5, alpha = 0.5) +   
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(FlowBoxMonth1b, outlierWY1b, outlierMonth1b)
```

 Merge the two data frames together
```{r}
Flow.outlier <- rbind(Flow.outlier.SHR, Flow.outlier.STTD)

summary(factor(Flow.outlier$Flow_Outlier))
```

26. Flowmeter Flags and New Flowdiff calculated

Cutoffs based on looking at plots and Upper limits calculated above. 

Flag   | Description                | Outlier                   | Flowdiff
Flag 3 | Highly Suspect             | Both                      | < 200 or >= 70000 or >= 50000 if inundation = FALSE
Flag 2 | Suspect                    | Tukey or MAD but not Both | 200-999 
Flag 1 | Acceptable                 |                           | 
No flag| Acceptable                 | None                      | 1000-59999

27. Add comment about what was flagged (FM for flowmeter-related)
28. Calculate replacement values where Flag_FM = 3 or Comment_SAMP = FM
  * Replacement values are median standardized flowdiff * settime based on groupings of Station, Flowmeter Speed, Water Year Class
  
```{r Flag-Replace-Flowmeters}
################## Flag Flowmeter data: Flag_QC1 ############################
# Mutate QC Flags 
# Conditions for Flag 3: Outlier = Both | Flowdiff < 200 | Flowdiff >= 50000 when no inundation | Flowdiff >= 70000
# Conditions for Flag 2: Outlier = Tukey | Outlier = MAD (but not Both) |Flowdiff between 200-999 
# Conditions for Flag 1: Acceptable data

# If Flag 3, calculate FlowdiffAdj = median.Flowdiff_s * SetTime

FlowmeterQAQC <- Flow.outlier %>%
  filter(!is.na(Flowdiff), Flow_modZ>=0.00000001) %>%
  mutate(Flag_FM = ifelse(Flow_Outlier=="Both" | Flowdiff < 200 | (Flowdiff >= 50000 & Inundation2 ==FALSE) | Flowdiff >= 70000 , 3, 
                          ifelse(Flow_Outlier %in% c("Tukey", "MAD") | Flowdiff < 1000, 2,
                                 "")),
         Comment_FM = ifelse(Flag_FM %in% c(2,3), "FM", ""),
         FlowdiffAdj = ifelse(Flag_FM == 3 | Comment_SAMP == "FM", median.Flowdiff_s * MeterSetTime, Flowdiff))
```

29. Summarize number of outliers flagged
```{r}
FlowmeterQAQC$Flag_FM <- ordered(FlowmeterQAQC$Flag_FM)
print("Flowmeter Outliers")
summary(factor(FlowmeterQAQC$Flag_FM))
print("Sampling Outliers")
summary(factor(FlowmeterQAQC$Flag_SAMP))
print("Lab Outliers")
summary(factor(FlowmeterQAQC$Flag_LAB))
```

30. Plot Changes and Flags
```{r}
outlierPreFlow <- ggplot(FlowmeterQAQC, aes(x = factor(Month), y = Flowdiff,  text = paste('Date: ', Date,'<br>Inundation: ', Inundation2, '<br>Set time: ', MeterSetTime), color = factor(Flag_FM))) + geom_point(size = 2.5, alpha = 0.5) +  scale_color_viridis(discrete = TRUE) + labs(title = "Flowmeter Pre-QC") +
  facet_wrap(~Station) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlierPreFlow)

outlierPostFlow <- ggplot(FlowmeterQAQC, aes(x = factor(Month), y = FlowdiffAdj, color = factor(Flag_FM))) + geom_point(size = 2.5, alpha = 0.5) + scale_color_viridis(discrete = TRUE)  + labs(title = "Flowmeter Post-QC") +
  facet_wrap(~Station) +   
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlierPostFlow)

grid.arrange(outlierPreFlow, outlierPostFlow, nrow = 2)
```

31. Plot before and after for SAMP
```{r}
outlierPreSamp <- ggplot(Flow.outlier, aes(x = factor(Month), y = Flowdiff, color = factor(Flag_SAMP))) + geom_point(size = 3, alpha = 0.8) + scale_color_viridis(discrete = TRUE) + labs(title = "Sampling Pre-QC") +
  facet_wrap(~Station) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlierPreSamp)

outlierPostSamp <- ggplot(FlowmeterQAQC, aes(x = factor(Month), y = FlowdiffAdj, color = factor(Flag_SAMP))) + geom_point(size = 3, alpha = 0.8) + scale_color_viridis(discrete = TRUE) + labs(title = "Sampling Post-QC") +
  facet_wrap(~Station) + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlierPostSamp)

grid.arrange(outlierPreSamp, outlierPostSamp, nrow = 2)
```


## QAQC: CPUE

32. Calculate Vol and CPUE, Vol_adj, CPUEAdj
```{r}
NetArea = 0.4572 * 0.25
FlowQAQC_CPUE <- FlowmeterQAQC %>%
  mutate(Volume = ifelse(FlowMeterSpeed == "Regular",
                         Flowdiff * 26873 * NetArea/999999,
                         ifelse(FlowMeterSpeed == "Low",
                                Flowdiff * 57560 * NetArea/999999, NA))) %>%
  filter(!is.na(Flowdiff)) %>%
  mutate(CPUE= round(Count/Volume,3))%>%
  
  mutate(VolumeAdj = ifelse(FlowMeterSpeed == "Regular",
                            FlowdiffAdj * 26873 * NetArea/999999,
                            ifelse(FlowMeterSpeed == "Low",
                                   FlowdiffAdj * 57560 * NetArea/999999, NA))) %>%
  mutate(CPUEAdj = round(Count/VolumeAdj,3))%>%
  arrange(Datetime)
```

33. Look at distribution of CPUE values
```{r}
FlowQAQC_CPUE$Month <- factor(FlowQAQC_CPUE$Month)
FlowQAQC_low <- filter(FlowQAQC_CPUE, CPUE < 10)
FlowQAQC_high <- filter(FlowQAQC_CPUE, CPUE > 10)

# Histogram of values
CPUEHist <- ggplot(FlowQAQC_CPUE, aes(CPUEAdj)) + geom_histogram(binwidth = 1) +
  facet_wrap(~Station) + theme_bw()

# SHR vs STTD boxplot
CPUEBox <- ggplot(FlowQAQC_CPUE) + geom_boxplot(aes(x = Station, y = CPUEAdj)) + theme_bw()

# SHR vs STTD boxplot by month
CPUEBoxMonth <- ggplot(FlowQAQC_CPUE) + geom_boxplot(aes(x = Month, y = CPUEAdj, color = Month)) + facet_wrap(~Station) + scale_color_viridis(discrete = TRUE) + theme_bw() 

grid.arrange(CPUEHist, CPUEBox)

# SHR vs STTD boxplot by year
CPUEBoxYear <- ggplot(FlowQAQC_CPUE) + geom_boxplot(aes(x = factor(WY), y = CPUEAdj)) + facet_grid(Station~.) + theme_bw() 
ggplotly(CPUEBoxYear)

grid.arrange(CPUEBoxMonth, CPUEBoxYear)
```

CPUE and Count Plots
```{r fig.width = 10, fig.height = 8}
# Species
ich_summary <-FlowQAQC_CPUE %>%
  filter(WY>1998) %>%
  group_by(Station, WYClass, ScientificName) %>%
  summarize(median.CPUE = median(CPUEAdj),
            max.CPUE = max(CPUEAdj),
            max.Count = max(Count),
            median.Count = median(Count),
            total.CPUE = sum(CPUEAdj),
            n = n()) %>%
  filter(n > 2)

# All CPUE
speciesCPUE <- ggplot(FlowQAQC_CPUE, aes(ScientificName, CPUEAdj)) + 
  geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
  scale_color_viridis(discrete=TRUE) +
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15)) 

ggplotly(speciesCPUE)

# Median CPUE
species2 <- ggplot(ich_summary, aes(ScientificName, median.CPUE)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

# MaxCPUE
species3 <- ggplot(ich_summary, aes(ScientificName, max.CPUE)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))


# Median Count
medCount <- ggplot(ich_summary, aes(ScientificName, median.Count)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

# Max Count
maxCount <- ggplot(ich_summary, aes(ScientificName, max.Count)) + geom_jitter(aes(col = WYClass)) + facet_grid(Station~.) + 
    scale_color_viridis(discrete=TRUE) + 
    theme_bw() + theme(axis.text = element_text(size = 12),
                      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

grid.arrange(species2, species3)
grid.arrange(medCount, maxCount)

# Which are the outliers?
ggplotly(maxCount)
```

34. Check on and change any known errors
```{r}
checkCounts <- filter(FlowQAQC_CPUE, Count>1000)
# Match up with datasheet 

checkCPUE <- filter(FlowQAQC_CPUE, CPUEAdj > 100)
checkCPUE2 <- filter(FlowQAQC_CPUE, is.na(CPUEAdj))
```

35. Look at summary table of adjusted CPUE values
```{r, eval = FALSE}
### Table of values
CPUE.adj.table <- FlowQAQC_CPUE %>%
  group_by(Station, FlowMeterSpeed, WYClass, ScientificName) %>%
  summarize(min.CPUE = min(CPUEAdj),
            median.CPUE = median(CPUEAdj),
            max.CPUE = max(CPUEAdj),
            n = n())

CPUE.adj.table %>%
    kbl() %>%
  kable_styling()
```


36. Determine and apply flags and cutoffs
    * In this case, used plots to decide to flag CPUE values > 70 

37. Add comment about what was flagged (Comment_CPUE = CPUE)

```{r}
CPUEQAQC <- FlowQAQC_CPUE %>%
  mutate(Flag_CPUE = ifelse(CPUEAdj>70, 2, "" ),
         Comment_CPUE = ifelse(CPUEAdj>70, "CPUE", ""))
```

38. Plot outliers
```{r}
outlier2a <- ggplot(CPUEQAQC, aes(x = Month, y = CPUE, color = Flag_FM)) + geom_point(size = 2.5, alpha = 0.5) + scale_color_viridis(discrete = "TRUE") + facet_wrap(~Station) + labs(title = "CPUE calculated with original Flowmeter values") + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

outlier2b <- ggplot(CPUEQAQC, aes(x = Month, y = CPUEAdj, color = Flag_CPUE)) + geom_point(size = 2.5, alpha = 0.5) + scale_color_viridis(discrete = "TRUE") + facet_wrap(~Station) +   labs(title = "CPUE calculated with adjusted Flowmeter values + CPUE flags") + 
  theme_bw() + theme(axis.text = element_text(size = 14),
                     axis.title = element_text(size =14),
                     legend.text = element_text(size = 14),
                     legend.title = element_text(size = 14), 
                     strip.text = element_text(size = 15))

ggplotly(outlier2b)

grid.arrange(outlier2a, outlier2b)
```

## Clean up display of dataset
39. Clean up order of variables
```{r}
# Rename some columns, remove unnecessary columns
ich_select <- CPUEQAQC %>%
  dplyr::rename(Inundation = Inundation2,
                FlagPhys = Flag_PQC,
                FlagSamp = Flag_SAMP,
                FlagLab = Flag_LAB,
                FlagFM = Flag_FM, 
                FlagCPUE = Flag_CPUE,
                CommentPhys = Comment_PQC,
                CommentSamp = Comment_SAMP,
                CommentLab = Comment_LAB,
                CommentFM = Comment_FM,
                CommentCPUE = Comment_CPUE) %>%
  select(c(event_id, Datetime, Station,
           WY, WYClass, Inundation,
           WeatherCode, Tide, MicrocystisVisualRank,
           WaterTemperature:Turbidity,
           ConditionCode,FieldComments, 
           MeterSetTime, FlowMeterSpeed, FlowMeterStart, FlowMeterEnd, Flowdiff, FlowdiffAdj, Volume,
           VolumeAdj,LabComments,SpeciesCode, CommonName, ScientificName, LifeStage, Count, CPUE, CPUEAdj, FlagPhys, 
           CommentPhys, FlagSamp, CommentSamp, FlagLab, CommentLab, FlagFM, CommentFM, FlagCPUE, CommentCPUE))
```


40. Merge flags and comments into one flag, one comment column, replace "" with NA
```{r}
empty_as_na <- function(x){
  if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
  ifelse(as.character(x)!="", x, NA)
}

ich_final <- ich_select %>%
  mutate(LifeStage = replace(LifeStage, LifeStage == "N/A", NA)) %>%
  mutate(across(c(FlagSamp:CommentCPUE), list(empty_as_na)) ) %>%
  select(-c(FlagSamp:CommentCPUE)) %>%
  rename(FlagSamp = FlagSamp_1, 
         FlagLab = FlagLab_1,
         FlagFM = FlagFM_1,
         FlagCPUE = FlagCPUE_1,
         CommentSamp = CommentSamp_1, 
         CommentLab = CommentLab_1,
         CommentFM = CommentFM_1,
         CommentCPUE = CommentCPUE_1) %>%
  mutate(QCFlags = paste(FlagSamp, " ", FlagLab, " ", FlagFM, " ", FlagCPUE)) %>%
  mutate(QCComments = paste(CommentSamp, " ", CommentLab, " ",CommentFM, " ",  CommentCPUE)) %>%
  relocate(FlowMeterSpeed, .before = "FlowMeterStart")

#look at data flags 3
threes <- ich_final%>%
  filter(grepl("3", QCFlags ))
```


41. Calculate proportion of data flagged
```{r}
## Amount replaced
print(paste0(round(nrow(threes)/nrow(ich_final)*100,3), "% highly suspicious"))
```


42. Write CSV file
```{r}
today = format(today(),"%Y%m%d")
write.csv(ich_final, paste0("R_write/Ichthyo/Ichthyo_data_", today, ".csv"), row.names = FALSE, quote = TRUE)
```

43. Make fish code crosswalk for IEP Fish Codes
```{r}
# fish species code crosswalk
SpeciesUpdate3 <- SpeciesUpdate2 %>%
  rename(CommonName1 = CommonName)

IEPFish <- IEPFish %>%
  rename(CommonName = `CommonName `)

Crosswalk <- full_join(SpeciesUpdate3, IEPFish, by = ("ScientificName")) %>%
  distinct() %>%
  filter(!is.na(SpeciesCode)) %>%
  select(-c(CommonName, Family)) %>%
  rename(CommonName = CommonName1) %>%
  relocate(SpeciesCode, IEPFishCode, CommonName, ScientificName)

write_csv(Crosswalk,"R_Write/Ichthyo/Crosswalk_fishcode.csv")
```

43. Write R Markdown File
